### 2017-05-24 renal1.rmd

---
title: "Renal 1"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
---

# Glomerulonephritis

3 cell populations in glomerulus = {endothelial, mesangial cells, epithelial cells (foot processes and parietal), } + 2 structural units = {basement membrane, mesangium}

Everything except for podocyte can regenerate

# Quiz (predominantly on UTIs)

-  Urine sample from 3/12 child: supra-pubic aspiration
-  Major mechanism causing stress incontinence: urethral hypermobility
-  Acute urinary retention requires urgen management
-  Non surgical Mx of BPH: $\alpha$-adrenergic-receptor antagonists
-  UTI w/ highly alkaline urine: *Proteus mirabilis*
-  Phimosis: foreskin cannot be retracted over the glans
-  Complicated UTI:
    -  First episode of acute cystitis in a sexually active 20 year old female.
	  -  *Acute pyelonephritis with fever in a 20 year old female.*
	  -  Urethritis in a 25 year old man who does not practice safe sex
	  -  Asymptomatic bacteriuria found on routine testing in a pregnant female
	  -  Bacteriuria associated with an indwelling catheter in a patient with neurogenic bladder
-  Lifetime prevalence of UTI in women: 50%
-  Trabeculation of bladder wall is hypertrophy
-  Medication associated with acute urinary retention: decongestant

# Random

## Targets of immune-mediated renal disease

-  Normal constituents of the glomerulus (e.g. Goodpastures syndrome; type II hypersensitivity)
-  Non-renal self antigens localised in globermuli => immune attack (e.g. SLE)
-  Exogenous antigens that localise in glomerular capillaries (e.g. PSGN)

# Main case

## Trigger 1

Allopurinol: xanthine oxidase inhibitor that reduces uric acid in the blood (treatment against gout)

What can cause urine to be red?

-  Blood (haemoglobin): test using urine microscopy/urinalysis = {ketones, villirubin / microscopy looks at RBCs, WBCs, epithelial cells {suggestive of non-midstream catch:not clean sample)} or urine dipstick test {protein (kidney disesae), pH, RBCs, leukocytes (WBCs), ketones(Diabetic Ketoacidosis; ketosis), specific gravity (osmolites/electrolytes in the urine), glucose (diabetes), bilirubin (cholestasis [nothing moving from gallbladder => spill into bloodstream]), nitrites (bacterial waste products)}
-  Beetroot/foods
-  Dehydration
-  Drugs/medications
-  Myoglobin
-  Menstruation
-  Porphyrins
-  Urobulinogen
-  Cholestasis: bile duct obstruction so it goes through the river (also has pale stools)

Haematuria: could be non disease cause such as trauma or exercise induced haematuria. >10E6/L is the defn cut-off.

Blood most likely to come from (trauma anywhere):

-  Glomerulus: leaky
-  Renal pelvis: due to calculi formation and trauma
-  Ureter: not so much as not very vascular
-  Bladder: cystitis, malignancy
-  Urethra: STDs (chlamydia clear discharge, gonorrhoea is purulent)

Pathological processes that can underlie haematuria:

-  Inflammation
-  Trauma
-  Infection
-  Malignancy
-  Vasculopathy
-  Autoimmune processes
-  Obstruction: of any hollow viscous
-  Necrosis/ischaemia: e.g. renal artery stenosis

Questions to ask

-  If pain => SOCRATES; ever had before
-  Fam Hx
-  Falls/trauma
-  Sexual Hx
-  Ever had before
-  Fam Hx

## Trigger 2

Triage: sick/unwell/well: briefly check red flags $\rightarrow$ send for some Ix

Focussed Ex: goes straight to the point/PC and deals with the less urgent stuff later, saves time

Hypertension $\rightarrow$ Hypertensive nephropathy (direct). More common is hypertension * (indirect) glomerular disease $\rightarrow$ haemoturia

Fenestrations in the foot processes of the podocytes. Allows small stuff and +ve ions.

DDx:

-  IgA nephropathy
-  Malignancy
-  Kidney stones
-  Infections
-  Trauma
-  Glomerulonephritis: glomerular disease
-  Coagulopathy such as haemophilia

Largest issue with low EGFR is hyperkalaemia: cardiac arrhythmias + death

Ix:

-  CHEM20: electrolytes, kidney, liver fn
-  Ultrasound KUB = {kidney, ureter, bladder}: see obstruction, fluid
    -  Contrast CT KUB is the only thing that shows about stones
-  Urinalysis
-  Urine cytology = looking for cells in urine

## Trigger 3

Membranous glomerulonephritis: common from Hep C

Dysmorphic red cells: have gone through the collecting tubules. Red cell casts: gone through the tubules  
Are pathognomic of glomerular disease (blood not coming from UTI or kidney stones)

Further Ix:

-  24h urine collection
-  Cystoscopy: puts camera into bladder, can argue against stones in the bladder, transitional cell carcinoma,...

## Trigger 4

Glomerulonephritis screen:

-  AMA: against autoimmune disease
-  dsDNA
-  Anti DNAase B: used as a screen for post infectious GN
-  Non nuclear auto-antibodies: Hep B, Hep C, and HIV
-  C3 C4 complement: found in post-infectious GN
-  ANCA: anti nuclear cytoplasm antibody
-  Anti-glomerular basement membrane: important for collagen disease

| Nephritic syndrome | Nephrotic syndrome |
| ------------------------ | --------------------------- |
| Blood in urine (RBC casts or dysmorphic cells) | Proteinuria >3.5g/24h |
| PHAROH = {proteinuria, haematuria, azotaemia, RBC casts (dysmorphic), oliguria, hypertension} | Proteinuria +++, hyperlipidaemia|
| Glomerulonephritis often a cause | ??? |

## Takeaways

-  Nephritic: PHAROH
-  Nephrotic: 3.5g/24h proteinuria AND oedema due to loss of hypoalbuminaemia. Children show it in the face.

Another way to look at it: glomerulonephropathies

-  Proliferative: too many cells (pink and purple) in glomerulus
-  Non-proliferative:

Biopsy look at it

-  H&E + microscopy
-  Immunofluorescence:
    -  Anti-GBM disease: the whole thing is lit up almost ribbon like
    -  IgA nephropathy: granular so lots of focal deposits
-  Electron microscopy

Focal vs segmental

-  Focal: whole glomerulus or part of glomerulus
-  Segmental: whole kidney, or just a part of a kidney

Watch the wikipedia [IgA nephropathy video](https://open.osmosis.org/). Tl;dr it is abnormally glycosylated IgA1 (malformed), don't get broken down and deposit in sites of filtration. Furthermore not recognised as self; breakdown in tolerance; immune complex deposition in mesangium. Immune response causes glomerular damage.

# Short case 1

## Trigger 1`

Glomerulonapathies:

-  Renal cause CVS: Decrease GFR => fluid retention => overloading the heart
-  CVS causing renal: vasculitis/renal artery stenosis => kidney failure => fluid retention => overloading the heart

Considering the timeframe: more like to have started a while ago and become protein ++++ in the past month.

Foamy protein = {proteins (such as albumin), increased force of urine, dehydration}

Glomerular damage => breach in the filtration system => heavy proteinuria.

Ultrasound KUB: doesn't say much about renal vasculature but looks at flow in ultrasound. Doppler shows blood flow through renal arteries.

Ix:

-  Urinalysis
-  24h urine collection: to see if it is actualyl nephrotic syndrome (by defn)
-  Bloods = {FBC, CRP, CHEM20, ddimer, lipid profile as }
-  Ultrasound KUB
-  Doppler ultrasound

UEC = urea, electrolytes, creatinine (EGFR)

UEG = UEC + glucose

LFTs = albumin

If need both LFTs and UEC then use CHEM20

## Trigger 2

DVT and kidney fn: nephrotic syndrome => losing proteins such as clotting factors (loses anti-thrombotic, loses pro-thrombotic but loses more anti-thrombotic due to the number?!) => hypercoagulable state

Hypertensive and kidney fn: decreased oncotic pressure

3 most common nephropathies for nephrotic syndrome

-  minimal change (children; bowmans space epithelial cells loses its negative charge, so not much morphological change)
-  Membranous glomerulonephritis:
-  Focal segmental glomerulosclerosis (FSGS):

## Trigger 3

Creatine clearance formula = Cockcroft-Gault Equation. Important as some drugs such as gentamicin have narrow therapeutic index.

Renal disease staging

1.  Normal eGFR, abno ultrasound
2.  eGFR > 60
3.  eGFR > 30
    a.  eGFR > 45
    b.  eGFR < 45
4.  eGFR > 15
5.  ESKD when eGFR < 15

Prognosis if you have no idea: young male, no comorbidities, proteinuria is alright

# Short case 2

## Trigger 1

-  Bilirubin breakdown clogged in liver, to avoid toxicity, body sends it ou bound to plasma in the urine (dark urine)
-  Assessing hydration status in children: mucous secretions, cap refill (centrally; push on chest 5s and see it refills in 2s)
-  Active urinary sedimen = {red cell casts, dysmorphic red cells}
-  GS = goodpastures syndrome
-  Henoch-Schonlein Purpura (HSP) = {Palpable purpura, abdo pain, arthralgia}

## Trigger 2

Hypertensive suggests renal disease.

Lymph nodes

-  Tender lymph node = infn/inflamm.
-  Non-tender lymph nodes = probs malignancy

Hypertension

-  Renal artery stenosis
-  Cushings
-  Renal disease
-  Malignancy: paraneoplastic effects (Conns, phaeochromocytoma)

Post infectious GN: circulating immune complexes form and clog up the glomeruli => inflamm response => damaged glomerular tufts => nephritic syndrome

-  Type III due to immune complex

Dysmorphic red cells pathognomic of glomerular disease