### 2017-05-24 renal3.rmd

---
title: "Renal III"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
---

Path

Dx of CKD requires:

-  Decline of kidney function for 3 months or more \textbf{AND}
-  Evidence of kidney damage (e.g. albuminuria or abnormal biopsy) textbf{OR}
-  GFR <60 mL/min/1.73 m2

Staging CKD

-  Stage 1: Kidney damage with normal or increased GFR (>90 mL/min/1.73 m2)
-  Stage 2: Mild reduction in GFR (60-89 mL/min/1.73 m2)
-  Stage 3: Moderate reduction in GFR (30-59 mL/min/1.73 m2)
-  Stage 4: Severe reduction in GFR (15-29 mL/min/1.73 m2)
-  Stage 5: Kidney failure (GFR <15 mL/min/1.73 m2 or dialysis)