### 2017-05-29 renal3.rmd
updated quiz

---
title: "Renal III"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
setwd("~/Documents/uni_work/4_sem_1/7111/lectures/15_renal_iii/readings")
```

# Readings

Dx of CKD requires:

-  Decline of kidney function for 3 months or more \textbf{AND}
-  Evidence of kidney damage (e.g. albuminuria or abnormal biopsy) \textbf{OR}
-  GFR <60 mL/min/1.73 m^2^

Staging CKD

-  Stage 1: Kidney damage with normal or increased GFR (>90 mL/min/1.73 m2)
-  Stage 2: Mild reduction in GFR (60-89 mL/min/1.73 m2)
-  Stage 3: Moderate reduction in GFR (30-59 mL/min/1.73 m2)
-  Stage 4: Severe reduction in GFR (15-29 mL/min/1.73 m2)
-  Stage 5: Kidney failure (GFR <15 mL/min/1.73 m2 or dialysis)

See Dan's autoregulation of kidney [**GIFT**](https://drive.google.com/drive/folders/0B0QeJC_rDeMIYjZMU2VfQVR2RU0)

# Quiz

-  Which of the following drugs must be used with particular caution in elderly patients with an existing low eGFR? 
    -  **Aspirin**
-  Which of the following indicates that kidney failure is chronic rather than acute? 
    -  Small kidneys
-  Pathophysiologic mechanism implicated in the development of diabetic nephropathy?
    -  Glomerulopathy due to glucose toxicity.
-  Epithelial cells on urine microscopy indicates contamination.
-  The commonest cause of death in severe chronic kidney disease is
    -   vascular disease
-  Greatest contributor to increase in CKD is diabetes mellitus NOT essential HTN/ageing population.

# Main case

![Was scribe](scibe-is-shit.jpg)

High anion gap is an indicator for metabolic acidosis.

Parenteral replacement is $\neg$oral (avoiding enteric system).

Hypocalcaemia $\rightarrow$ hyperPTH and hypophosphataemia + hyperphosphaturia (lack of PO4 reabsorption in the nephrons).

# Short case 1

## Trigger 1

35M presenting with increasing lethargy and weight loss. No fevers, night sweats, or pain.

Past Hx: 12 years old Dx with T1D.

Pathophysiology of diabetic nephropathy: glycosylation of basement membrane $\rightarrow$ leaky $\rightarrow$ proteins flow out $\rightarrow$ thickening of arteriole (mainly efferent) [vasoconstriction] $\rightarrow$ decreases blood flow to peritubular capillaries $\rightarrow$ they start dying.

-  ACEi dilate efferent which renoprotects against this efferent constriction at the cost of decreased GFR
-  Scarring due to HTN, glomerulopathy usually leads to atrophy due to fibrosis and it shrivels up, whereas due to diabetes has normal sized kidneys: sugar gets deposited in the kidney with sugar crystals and glycosylation (shit kidneys, low GFR, normal sized)

Presentation for CKD does not have obvious symptoms.

## Trigger 2

Antimicrobials

-  Co-trimoxazole = trimethoprim + sulphamexazole (folic acid inhibitors at different parts in the pathway; synergistic). General prophylaxis against bacteria. Good against gram(+) such as Staph.
-  Valganciclovir = antiviral. Targeting CMV
-  Amphotericin lozenges = antifungal

HLA-DR3,4 is *extremely* common with T1D.

Mechanisms of rejection

1.  HLA incompatibility
2.  ABO: blood type
3.  Rh: rhesus factor
4.  Cross-matching: check if recipient has any miscellaneous antibodies (anti-{Kell, Duffy, White} are some common ones)

Immunosuppresive agents:

-  Calcineurin inhibitors: tacrolimus and cyclosporin act and prevent IL-2 formation $\rightarrow$ no T cells
-  Antimetabolites: mycophenalate prevent B and T lymphocyte production
-  IL-2 R antag: basilixamab

Side effects: opportunistic infections, cushings, malignancy, osteoporosis (w/ glucocorticoids)

# Short case 2

## Trigger 1

42M severe right flank pain and is now all the way to his groin. Pain is coming in waves and came in 1 hour ago.

DDx: renal colic caused by kidney stones (aka nephrolithiasis), biliary colic (gall stones biliary tree), pyelonephritis, PKD $\rightarrow$ cyst rupture, (Female can be ovarian stuff), renal capsule stretched, malignancy

Colicky pain in hollow viscous with peristalsis.

Pain radiation: referred pain due to nerves such as iliohypogastric

Pathophysiology of HTN: build up of fluid $\rightarrow$ Down GFR $\rightarrow$ RAAS to increase BP. HTN is a risk factor for kidney stones

Ix to order:

-  Xray KUB: Ca stones (80-90%) are visible since it is a metal
-  Ultrasound KUB: visualise hydronephrosis (swollen proximal ureter and thin distal ureter) due to an obstruction, can see calcification, see if kidney is swollen, bladder = {stones in there?, vesicoureteral junction reflux}
-  CT KUB: if ultrasound is not diagnostic
-  CHEM20 can view the PO4, Ca2+, Mg2+, uric acid  which may cause the stone | measure eGFR which can indicate for surgery
-  Urine MCS: bacteriuria, casts/dysmorphic red cells, pyelonephritis, septic cause, maybe see crystals (oxalate)
-  FBC: white cells can indicate pyelonephritis
-  Urinalysis: proteinuria or haematuria, nitrites pathognomic of UTI (which includes pyelonephritis)

## Trigger 2

Kidney stones often made up of: Ca, oxlate, NH4+, PO4, cysteine, nitrogen compounds

Risk factors urolithiasis: Hx of gout, foods high in oxalate (beetroot, strawberries), alkaline urine (risk for uric acid stones), diabetes, immobility (stasis; such as from obstructive disease like **PKD**), males, fam Hx, drugs, dehydration

Stones can lead to infection, hydronephrosis, decreased GFR, AKI $\rightarrow$ CKD, shitload of pain, rupture of ureter

Lifestyle Mx: hydration, change diet (less of some shit, less beer), exercise

Tx:

-  Medical: lithotripsy can break the stone down if it is small enough (<2mm), alpha blockers or CCB (?!; is possible)
-  Surgery: urological surgery (by putting a stent in and allowing it to pass through

Key factor: if the stone is >10mm then operation. Not then try medical treatment