### 2017-06-02 pathreview.rmd
I'm dead now

# Hints

-  Learn path from tutes and go there to view it
-  Learn clinical vignette, pathophys processes, sequelae, risk factors, associations, etc.
-  Histo: learn fn of the tissues you are studying

# Micro

Clostridium: rod | Strep are NOT cat(+) | $\beta$-haemolytic are MRSA

Beta is complete, alpha is partial, gamma is almost non-existent

# Atherosclerosis

High yield. Location is important too. SO abdo aorta is most likely to occur.

Thrombus from rubture in plaque | Embolus: plaque moves around

Pathophys: smooth muscle from tunica media. Exposure of subendothelial collagen => platelet aggregation and activation

## Histo

Foam cell: clear, circular, and has a nucleus. Chloesterol clefts are the sword looking imprints. Smooth muscle does exist.

Plaque ruptured => traps RBCs. Pale area: fibrous-looking, contains RBCs and vessels = granulation tissue. Type III collagen is deposited => switches to Type I.

# Arteriosclerosis

3 types. Hyaline in general refers to pink proteinaceous material. Learn about AD-PKD since that is a fairly common Qs.

# Aneurysms

Circle of Willis; no internal elastic lamina => saccular (berry) aneurysms

Learn ehlers-danlos

# Aortic dissection

Blood goes from lumen into intima. Elastic tissue not holding up. Vasa vasorum are vessels that supply other vessels.

## Histopath

Intima is the top, because vasa vasorum is visible at the bottom.

# DVT

Phlebothrombosis are thrombosis in veins | Has lines of zahn which are aggregates of fibrin and platelets, means it is ante-mortem (pre-death).

# MI

-  B = blue cells are neutrophils so it is probably early hours
-  C = nuclei starting to disappear so probably still early
-  D = still massive amounts of neutrophilic shit everywhere so probably day 2 or 3
-  E =  fibroblasts and vessels = granulation tissue

Acute Inflamm: first cell is neutrophils, after day 2 macrophages come in along with granulation tissue, deposit type III collagen, then replaced by type I collagen

NTK: what time points are MI shit happening (histo)

Evolution: coagulative necrosis is when nuclei start to fade. Early granulation tissue is Type III collagen which can still rupture.

Pathology: aneurysm in ventricle

Know Virchow's Triad

# PE

E.g. Someone completed surgery and had a PE. List some risk factors:.:

# Infarcts

Red infarcts: haemorrhagic => due to collaterals. Lungs, liver, intestines

Pale: => end arteries such as heart, kidney, spleen

# Infective endocarditis

Gram stain pink = gram neg

# RHD

Pathognomic are aschoff bodies. Not often zoomed in either. Causative organ Strep pyogenes: E (purple cocci). D is probs Staph. A is enteroccosus?..

Tammy likes molecular mimicry.

Learn Jones criteria

Histo: aschoff nodules have caterpillar nuclei

Sequelae of viral myocarditis (a short case): dilated cardiomegaly | Probs should learn

# Asthma

2 features: mucous plugging, dilation of bronchial wall. UQ likes asthma. Reversible bronchoconstrictionHisto changes: mucous producing cells are goblet cells.

## Histo

Really pink granulated cells are eosinophils in the centre of the bronchus. Basement membrane is thickened (too pink). Cells have infiltrated (blue) are lymphocytes. "See blue cells in inflammatory setting => LYMPHOCYTES".

Drugs and pathophysiology are high yield. Type I hypersensitivity <= Tammy likes this. Th2 subtype is involved in sensitisation. ACh in the lung => bronchoconstriction.

# Emphysema

Bullae => pneumothorax

Types: centriacinar = mostly restricted to resp bronchioles | panacinar: cna consume whole alveolar space. Smoke causes centriacinar as it gets stuck there and can't progress further (usually upper lobe). Panacinar is seen in AAT.

Know lung fn tests:

# Bronchiectasis

Top lobe looks good, bottom has dilation of bronchioles. Is obstructive. Resp epithelium = pseudostratified ciliated columnar.

# Restrictive disease: silicosis

Sarcoidosis would appear earlier than 60.

Fibroblasts come in and try to contain the silica => produce collagen. Nodules build up and cause inflammation.

# Pneumoconiosis

Simple CWP no cancer. COmplicated CWP: higher chance of cancer. Caplan's syndrome

# Asbestosis

## Histo

Extremely thick pleura filled with connective tissue (fibrosis; collagen) and there is some inflammatory infiltrate (lymphocytes).

# Granulomatous inflamm

Activation of macro by Th1 cells secreting IL12. Eat everything then form a giant cell.

# Sarcoidosis

Non-caseating granuloma + hilar lymphadenopathy. Macro: hypercalcaemia due to hyperphosphate?!

# Pneumonia

Greay hepatisation: full of neutrophils exudate and fibrin. Lobar more acute presentation. Learn the stages of pneumonia.

Microbiology in short cases

Gram stain | Shape | Feature
E.coli, strep, Learn the atypicals that cause interstitial pneumonia.

# Neoplasms

HIGH Yield

Keratin pearl = squamous cell.

Learn what metaplasia is UQ likes that.

# Lung tumours

## Adenocarcinoma

Glandular on histo. Scar see collagen.

Paraneoplastic: lambert blocks presynaptic neuron => flaccid paralysis.

## Path/histo

Benign = well circumscribed mass.

# Kidney

# Acute pyelonephritis

# Renal cell carcinoma

carcinoma in-situ: has not yet invaded basement membrane

Chromophobe stains pinkish. Need to know clear cell one

VHL Genetics: signal to produce EPO is hypoxia.

Sarcomas: tumour of connective tissue (fat,...) spread through blood

Carcinomas generally spread lymphatically.

## Wilms tumour

UNdifferentiated embryological. Genes WT1 and WT2.