### 2017-06-02 CVS5.rmd

---
title: "CVS 5"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
setwd("/home/khoo/Documents/uni_work/4_sem_1/7111/lectures/8_cvs_v/readings")
knitr::opts_chunk$set(echo = FALSE)
```

# Quiz

## Medications

-  Dobutamine: Is a $\beta_{1}$ agonist so it will increase myocyte contractility and HR.
-  Diltiazem: reduces afterload and hence myocardial work.
-  $\beta$-blockers used in HF are only: metoprolol, bisprolol, and carvedilol

## Acute Rheumatic Fever

Symptoms:

-  Subcutaneous nodules
-  Carditis
-  Polyarthritis
-  Chorea
-  **Pathognomic**: aschoff bodies

## Aschoff nodules

Aschoff nodules are composed of T lymphocytes (not B lymphocytes) and multinucleated giant cells. Not seen until the subacute or chronic phase of rheumatic carditis.

## Ventricular assist device

Given to advanced HF pt with good prognosis.

## Complications of advanced heart failure

-  Hyponatremia
-  Impaired liver fn
-  Thromboembolism
-  Arrhythmias
-  NOT \textbf{hyper}natremia

## Pericardial tap

aka pericardiocentesis: needle and catheter remove fluid from pericardium.

Pericardial fluid has an exponential pressure volume relationship.

# Readings

## CXR

```{r image, out.width="500px"}
knitr::include_graphics("cxr.png")
```

[\textbf{\textit{url}}](http://www.radiologymasterclass.co.uk/tutorials/chest/chest_pathology/chest_pathology_page8#top_1st_img)

-  Septal lines are a specific sign of interstitial oedema in suspected heart failure. If not suspicious of HF then conditions causing lymphatic obstruction can be a cause.
-  Alveolar oedema: general shadowing of the lungs. Dense airspace due to fluid filling alveoli and small airways radiating from the hilar regions.
-  LV aneurysm: it just sticks out on the side
-  Consolidation of the lingula: poorly defined L heart border due to consolidation in the adjacent lingula.

# Random path terms

-  Petechiae: pinpoint areas of haemorrhage on skin or mucous membranes reflecting bleeding from capillaries.
-  Purpura: haemorrhages of skin or mucous membranes larger than petechiae.
-  Suppuration: formation/discharge of pus
-  Empyema: pus in cavity or hollow organism
-  Coagulative necrosis (heart): 3 characteristics = {loss of nuclei, increased eosinophilia of the cytoplasm, maintenance of the remnants of the cell outline}
-  Liquefactive necrosis: complete dissolution of cells and other structures
-  Fat necrosis: seen in adipocytes as a result of trauma or release of lipases
-  Karryorrhexis: fragmentation of the nucleus seen in necrotic and apoptotic cells
-  Exudate: protein rich fluid resulting from the selective extravasation of plasma in response to changes in vascular permeability during inflammation

# POW Week 8

-  Measure someones BP and it is 160/90. Not hypertension unless it is *measured again*
-  Cardiac hypertrophy (not hyperplasia as mature myocytes cannot divide) is a response to increased BP
-  LVH response to increase in systemic BP; RVH response to increase pulmonary BP
-  Types of hypertrophy
    -  Concentric: increase pressure causes thickening of walls but lumen roughly the same
    -  Eccentric: volume overload (increase in end diastolic volume) causes increase in lumen size whilst walls may/not hypertrophy
-  Causes of cardiac hypertrophy
    -  MI: loss of some of the cardiac muscle causes remaining muscle to adapt to increased workload
    -  Hypertrophic cardiomyopathy: genetic mutation causing thickening
    -  Cor pulmonale: pulmonary arterial hypertension leading to RV failure
    -  Aortic stenosis: valvular disease causing increased pressure load on RV
-  Outcomes of LVH:
    -  Increased tissue pressure in wall, harder for blood to perfuse during diastole
    -  Increased muscle mass requires more oxygen
    -  Combination of above 2 predisposes to ischaemic conditions

# Random notes on ECG

-  AF loses *atrial kick* and is a dribbly pattern
-  AFlutter is sawtooth baseline and occurs after the formation of a reentrant circuit
-  SVT: very fast rate with narrow QRS and
-  Second degree heart block 2:1 means that every alternative P wave doesn't conduct a QRS
-  Junctional (escape) rhythm: pacemaker from AV node or bundle of His. Absent or inverted P waves in lead II. Due to not much atrial contraction there is also loss of atrial kick
-  Accelerated junctional: loses P and T waves
-  Junctional tachy: (my own words) shark tooth QRS
-  Premature ventricular complex (PVC): increase in autorhythmicity or reentry
-  If P waves abnormal but still in sinus rhythm then it is NSR with PAC (premature atrial complex)

# CBL main - waking up breathless

## Trigger 1

68 F: Dyspnoea, orthopnoea (dysponea whilst being straight), PND, panicky, easy fatigueability, reduced exercise tolerance (e.g. can't push her buggy around anymore), chest tightness

-  SOB at night
    -  PND: peripheral venous blood redistributed to lungs. Increased venous return $\rightarrow$ more blood in pulmonary circulation $\rightarrow$ increase preload $\rightarrow$ LV can't keep up pumping into systemic circulation (LV failure) $\rightarrow$ congestion in the lungs $\rightarrow$ transient oedema
    -  Difference between PND and orthopnea: **PND**: go to sleep easily but then they wake up. Often decompensated orthopnea vs **Orthopnea**: difficulty falling asleep as SOB as soon as lying down. Ask by "do you get SOB when trying to sleep"
    -  COPD, asthma, OSA
-  SOBOE: typically symptomatic of IHD (CAD is just related to state of arteries).
-  **Chest pain DDx $\Leftarrow$ MEMORISE**:
    -  Cardiovascular - pericarditis, IHD
    -  Respiratory: PE, pneumonia $\rightarrow$ pleuritis
    -  MSK: costochondritis (Tietze syndrome: palpation causes pain),
    -  GIT: gastritis $\pm$ ulceration (peptic ulcer), reflex, oesophageal spasm
    -  NS: thoracic nerve root irritation (compression of nerve root), herpes zoster (shingles)
    -  $\psi$: anxiety
-  Tiredness: infection, anxiety, depression, anaemia, cancer, IHD, vavlular disease, thyroid disease
-  Hypoxaemia $\rightarrow$ sympathetic drive $\rightarrow$ panic attack (/disorder: repeated unpredictable panic attacks or spending lots of time about the impending panic attacks. Need to treat AND educate patients)
-  What do next? Hx (HPC,  systems review, meds, surgical Hx, family Hx, Allergies) $\rightarrow$ Ex $\rightarrow$ blah blah

## Trigger II

-  Acute on chronic: chronic disease that suddenly becomes more acute
-  Chest discomfort/**anything**: cardiac until proven otherwise - do ECG. Something that can't be missed.
-  Narrow Dx to cardiovascular: due to family Hx of vascular disease, hypertension. Probs not respiratory: has no palpitations, fever, cough or wheeze, does not smoke.
-  Calcium channel blockers
    -  Dihydropyridines: amlodipine, nifedipine $\Leftarrow$ ACTS PERIPHERALLY
    -  Non-dihydropyridines: verapamil & diltiazem $\Leftarrow$ acts centrally. Slows conduction through the AV node, negative inotropic and chronotropic activity.
-  Potential Dx: heart failure $\rightarrow$ crackles of lung bases, pitting oedema, elevated JVP, displaced apex beat, S3 (increased V filling pressures), might feel heaves
-  Pulsatile liver is pathognomic of tricuspid regurg

## Trigger III - where did her HF come from? Cardiomyopathy

-  Turbulent flow in the femoral arteries: stenosis, atherosclerosis
-  Bibasal crackles: APO $\rightarrow$ oedema fluid could be pus (pneumonia), blood, interstitial fluid. Lots of things that cause it, but in this pt other symptoms suggests oedema fluid
-  AV nipping and silver wiring: signs of hypertensive retinopathy (signs of hypertension in the eye)
-  She needs treatment but not that urgent as sats O~2~ is not that bad
-  Ix: ECG, bloods (FBC, CHEM20 - kidney(could lead to fluid retention $\rightarrow$ overload)/liver(late sign)/electrolytes(retention of Na due to RAAS leads to water retention + hyper/hypo-kalaemia lead to arrhythmias), **BNP** (peptide released from ventricles; use if not sure about HF: high = HF), troponin, lipid profile), CXR, echo
-  Hypercholesterolaemia $\subset$ Dyslipidaemia (too much or too little anything)

## Trigger IV

**ECG**:

-  Right axis deviation suggests RBBB
-  Path Q waves in V1, V2, V3: previous anterior MI
-  Lead III: the notching in the QRS complex suggests BBB
-  S waves normal in I and II
-  PR interval long = first degree heart block
-  Broad QRS: either ventricular or atrial with aberrant conduction through AV node (this is more likely since she has BBB)
-  V1: T wave is the inversion and sort of cuts the S wave off (change in gradient)

**Blood test results**:  Has dyslipidaemia. Thyroid fn measured as TH affects HR and contractility by the SNS. BNP is a specific test: can help see if treatment is working.

**CXR** signs of HF: may not see all 5, but quick to identify

-  A: alveolar oedema (bat wings) - cloudiness near the hilum
-  B: kerley B lines: septal thickening
-  C: Cardomegaly (Cardio thoracic ratio)
-  D: diverted upper vasculature: blood gets shunted to the top of the lungs since there is too much fluid. Basically see what you usually see inferior, see it superior
-  E: effusion: blunting of the costophrenic angles due to fluid build up

Pansystolic murmur means not surprising that there is tricuspid regurg (radiates to axilla?)

Echo gives valves and LV EF

Silent MI: most common in diabetics, can happen in women

## Trigger V

-  All Ca2+ channel blockers are contraindicated in HF as it decreases contractility of the heart.
-  Frusemide keeps Na and loses K, so give with K supplement (Span K)
-  Don't start medications that decrease work of the heart until *euvolaemic* - oedema is gone.
-  ACE inhibitor: dry cough due to bradykinin. Decreases BP and also helps with cardiac myocyte remodelling after an infarct.
-  Aspirin: low dose acts as antiplatelet agent by working on ADP.

## Trigger VI

-  Myocardial perfusion scan: uses radioactive contrast and get stressed through drugs (usually dobutamine which is a sympathetimimetic)
-  Multivessel disease means it is good for CABG. If just a single vessel then it is not worth
-  Stress testing is to examine perfusion. Usually stress echo $\rightarrow$ something not moving $\rightarrow$ angiogram
-  PTCA = percutaneous transluminal coronary angioplasty is an alternative to the heart surgery
-  Bypass surgery can result in death, haemorraghe, clotting, infn
-  LV failure can lead to RV failure leading to congestive HF.

[\textbf{\textit{Good ecg website}}](http://www.skillstat.com/tools/ecg-simulator)

# Midsem quizzes

## CVS anat and physiology

-  If SA node fails then AV node takes over. Internodal pathways don't do anything. 
-  Long saphenous vein is *superficial*
-  Aortic valve has 3 cusps; right coronary, left coronary, and *posterior*/non-coronary cusps.
-  LAD is most prone to obstruction
-  Paradoxical embolism: taken from venous system to arterial (such as through patent foramen ovale)
-  VSD is a L$\rightarrow$R shunt; tricuspid atresia is a R$\rightarrow$L shunt - an ASD *and* VSD must be present for blood to be pumped due to the RV atrophy. Hence it is a shunt.
-  Azygous vein is on the *right* and *posterior* thoracic wall
-  Opthalmic artery is a branch of the *internal carotid*
-  **$\Uparrow$Pulse pressure caused by:
    -  $\Downarrow$aortic compliance
    -  $\Uparrow$stroke volume
    -  Hypertension

## Embryo

-  Just after endocardial tubes have fused this is the order of blood flow: sinus venosus, primordial atrium, primordial ventricle, bulbus cordis, aortic sac
-  Development of foramen ovale: common atrium starts by growing septum primum downwards $\rightarrow$ ostium primum randomly forms and disappears $\rightarrow$ above an ostium secundum forms $\rightarrow$ septum secundum forms which is slightly offset forming a one-way valve.
    -  Tl;dr Septum primum, ostium secundum, foramen ovale, septum secundum
-  Ductus arteriosus closes due to
    -  Increased arterial oxygenation
    -  Decreased pulmonary vascular resistance
    -  $\Downarrow$Prostaglandin E~2~

## Reperfusion injury

-  O~2~ is incompletely reduced by damaged mt $\rightarrow$ $\Uparrow$O~2~ free radicals
-  Influx of activated leukocytes increases inflamm damage

# Short case 1

## Trigger 1

-  Cardiac pain (ischaemic) is not usually sharp (instead crushing, heavy, dull).
-  Chest pain to ignore MI: 3 P's = {palpable, pleuritic (breathing makes it worse), positional (pain changes on changing position)}
-  Positional chest pain likely pericarditis: pericarditis is inflammation of pericardium, moving around causes friction which irritates the already inflammed pericardium, gravity puts pressure on the pericardium.
-  Chest pain is usually urgent: DDx = cardiovascular, resp, neuro, musculoskeletal
-  What other info is needed: SOCRATES, PMHx (could be used to rule out oesophageal reflux,...), PSHx, medications, allergies

## Trigger 2

-  ECG
    -  Widespread ST elevation can be found in pericarditis
    -  Look for deep S waves in V1, V2
-  Pulsus paradoxus: drop in pressure during inspiration where you'd usually expect an increase due to increased venous return.
    -  Pericardial effusion is fluid around the heart
    -  Pericardium filled with fluid is known as tamponade.
-  Becks triad to Dx tamponade: hypotension, raised JVP, soft heart sounds
-  Ix: bloods (FBC for anaemic, CRP/ESR, troponin, cultures, CHEM20 especially because have to check renal/kidney fn before giving them medication, coagulation panel in case he is just bleeding into it, BNP in case HF)
-  Pericardial tap to drain the fluid

## Trigger 3

-  CXR: round heart is a sign of pericardial effusion
-  3 layers of the heart: endocardium, myocardium, epicardium
-  Troponin leak from the myocytes adjacent to the epicardium because of inflammation
-  BNP raised: decreased aortic BP (hypotension) causes $\Uparrow$BNP
-  ISBAR = {identify, situation, background, assessment, recommendation}
    -  "On Ex"
    -  "Ix shows...typical balloon sign of pericarditis. In addition, echo performed shows 2cm pericardial fluid" $\leftarrow$ which is the most important thing rather than just LVH etc.

## Trigger 4

-  Myopathy: diseased myocardium
-  Viral myocarditis $\rightarrow$ $\Uparrow$troponin $\rightarrow$ pericardial effusion
-  Hyperaemia: increased blood supply
-  Oedema: increased interstitial fluid
-  Rheumatic heart disease: Group A Strep pharyngitis $\rightarrow$ autoimmune response
    -  Jones criteria

# Short case 2

## Trigger 1

-  SOB at rest: NYHA 4
-  Tricuspid regurg $\rightarrow$ pulsatile liver
-  Lung disease $\rightarrow$ pulmonary hypertension $\rightarrow$ RHF/heart dysfn
    -  known as *cor pulmonale* (heart disease due to lung pathology)
-  Aspects of physical Ex and why: CVS Ex: focus on the heart and lower limbs.
    -  Elevated JVP
    -  Palpation of the chest: parasternal heaves for [L,R]VH, lateral displaced apex beat due to LVH
    -  Auscultation: valvulopathy as a cause of this, extra heart sounds (S3 due to HF)
    -  Lower limb Ex: oedema (pitting)
    -  lung auscultation(?): bibasal crackles suggestive of fluid from____(pneumonia)

## Trigger 2

-  Ex for work of breathing
-  Loud P2 from pulmonary hypertension
-  Ix:
    -  ABG
    -  Bloods (FBC for [Hb] to *exclude anaemia*, CHEM20 ensure kidney+liver fn before giving drugs, COAG panel looking for thrombosis, ESR/CRP: see if infn is causing),
    -  ECG: hypertrophy of ventricles in HF
    -  CXR: signs of HF, ABCDE={alveolar oedema, kerley \textbf{b}-lines, cardiomegaly, dilated upper vessels, pleural \textbf{e}ffusion}}
    -  Echo (EF)

## Trigger 3

-  Haematocrit is volume percentage of RBCs in the blood.
-  Slightly elevated red cell count (RCC) and haematocrit is a compensatory mechanism as a result of chronic hypoxaemia
-  COPD with high CO~2~ (CO~2~ retainers): the chemoceptor changes from peripheral CO~2~ detector changes to central receptors which are *dependant on O~2~*. Therefore giving O~2~ will decrease respiratory drive $\rightarrow$ death.
-  ECG
    -  Right axis deviation: high V1
    -  Inverted T-waves normal in V1 (upside down), whereas not normal in V2, V3 due to RV strain (ischaemia)
-  Mx
    -  Decrease peripheral oedema: *frusemide* to get the fluid off the lungs (alveolar oedema) and legs $\rightarrow$ euvolaemic
    -  Oxygen therapy: contraindicated with smoking
    -  Pulmonary and cardiac rehab that improves the muscles