### 2017-06-02 I&I.rmd

---
title: "I&I 1 & 2"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: 
  pdf_document:
    keep_tex: true
---

```{r setup, include=FALSE}
setwd("/home/khoo/Documents/uni_work/4_sem_1/7111/lectures/3_i&i_ii/readings")
```

# Bacteria

-  Gram(-) rod: *E. coli*
-  Gram(-) coccus (round): *Neisseria meningiditis*
-  Gram(-) rod after burn with green and foul smelling exudate: *Pseudomonas aeruginosa*
-  Gram(+) bacili can form *spores* in order to resist environmental factors

# Drugs

-  Efficacy: maximal response
-  Potency: how fast it starts working
-  In the presence of competitive antagonist, the agonist dose-response curve is:  shifted to the right without changing slope or maximum effect
-  Selective toxicity is NOT: Clavulanic acid acting as a target for beta-lactamase

## Receptor

-  Desensitisation: decrease in potency (curve shifts right) usually due to partial uncoupling of receptor from second messenger
-  Down-regulation: decrease in efficacy (maximal effect) usually due to receptor internalisation
-  Not often seen for drugs with spare receptors

## Anti-bacterials

-  Interfere with folic acid synthesis so DNA synthesis cannot occur: sulfonamides (e.g. trimethoprim)
-  $\beta$-lactams: attach to penicillin-binding proteins on bacteria and interfere with peptidoglycan (cell wall) synthesis. Also cause lysis
    -  E.g. Penicillins, cephalosporins, glycopeptides (vancomycin)
-  Interfering with bacterial protein synthesis: 
    -  50S: macrolides (erythromycin)
    -  30S: tetracyclines (used for intracellular bacteria such as Chlamydia),  aminoglycosides (streptomycin, gentamicin, neomycin against gram(-) aerobes)
-  Affecting topoisomerase: quinolones

## Antiviral

Gancyclovir is against CMV. Acyclovir prevents nucleic acid synthesis. Enfuvirtade used in HIV: inhibits viral cell penetration/fusion.

## Anti-protozoan

Malaria $\in$ {Protozoa}. Therefore quinine and chloroquinine are used to target the schizont stage. Metronidazole inhibits nucleic acid synthesis and is also antibacterial.

## Anti-helminth

Its location means that Combantrin (non-spec depolariser) is used with poor absorption through GIT.

## Anti-fungals

Fluconazole and clotrimazole inhibit ergosterol production used against Candidiasis. Amphotericin used for life threatening (probably deep mycoses).

# Anti inflamm pharm

-  SAIDS: glucocorticoids indirectly inhibit phospholipase A by upregulating TXN of annexin-1
-  NSAIDS: inhibit COX which prevents synthesis of PG~E2~ which is a pyrogen. $\therefore$ they are anti-{inflammatory, pyretic} and analgesic. E.g. aspirin, ibuprofen, celecoxib, 
-  Paracetemol: NOT an NSAID. MOA unknown.

# Immunology stuff

\textbf{M}ost \textbf{D}octors \textbf{G}et \textbf{E}xtra \textbf{A}ss - IgA is in **secreted fluids**

\textbf{Hot} \textbf{T}-\textbf{B}one st\textbf{EAK}

-  IL-1: increases temperature (fever) making it an endogenous pyrogen. It is secreted from *macrophages* and also increases adhesion factors on the endothelial wall during inflammation.
-  IL-2: T cell stimulator
-  IL-3: B cell stimulator
-  IL-4: IgE (class switching from IgG) and B cell growth. Secreted from T~H~2 cells.
-  IL-5: IgA and eosinophils. Also secreted from T~H~2 cells.

T~H~1 secrete: (leading to macrophage mediated chronic inflammatory response)

-  IFN-$\gamma$: MOST SIGNIFICANT activator of macrophages along with IL-2(?)
-  TNF-$\alpha$: helps macrophages and is important in autoimmune disorders
-  IL-2: T cell stimulator

T~H~2 secrete: (main effectors are eosinophils, basophils, mast cells, and B cells)

-  IL-4: IgE + B-cell growth
-  IL-5: IgA
-  IL-6: Ac(K)ute phase protein synthesis
-  Supports B cell antibody responses

T~H~17: secretes proinflamm cytokines such as IL-22 which stimulate production of bacteriocidal and fungicidal proteins by epithelial cells.

Types of hypersensitivities: **ACID**

-  Type I: \textbf{A}naphylactic and atopic: mediated by basophils and mast cells
-  Type II: \textbf{C}ytotoxic (antibody mediated)
-  Type III: \textbf{I}mmune complex
-  Type IV: \textbf{D}elayed (Cell-mediated)

Fever

-  Temperature is raised by pyrogens: exogenous and endogenous. Exogenous includes LPS from gram(-) bacteria. This triggers response leading to IL-1 secretion which migrates to the hypothalamus signalling for increase in body temperature.
-  Types: Intermittent (returns to normal and goes back), remittent (varies but never goes to normal),
-  Pyrogens: IL-1, IL-6, TNF-$\alpha$

PAMPs

-  LPS is found in gram(-) bacteria which is detected by TLR-4
-  Lipoteichoic acid from gram(+), peptidoglycan, viral nucleic acid (such as dsRNA) are detected by TLR-3.
-  Microorganism binding to PAMP-R leads to cytokine activation.

Random notes

-  B cell receptor is bivalent whilst TCR is NOT
-  Most important ion for determining RMP is K^+^
-  Volume of distribution is most important in determining loading dose of a drug
-  Neutrophils are necessary for umbilical cord separation.
-  NK cells are important *early* and cytotoxic t cells (CD8) are important *late* in the course of a viral infection.
-  During acute inflammation, blood vessels in the affected region of the body dilate leading to an overall increase in blood flow: most likely to be present is Increased adhesion of leukocytes to endothelial cells

Activation of complement leads to:

-  **Phagocytosis**: by opsonising antigens. C3b is most important here
-  **Inflammation**: attracts macrophages and neutrophils
-  **Membrane attack**: directly rupturing cell membrane of foreign pathogens
-  Does NOT activate Th cells

Macrophages:

-  Phagocytose bacteria
-  Present antigen on MHCII molecules
-  Recruit neutrophils to a site of infection
-  Responsible for controlling release of cytokines
-  Reside in tissues ($\neg$ in lymph nodes)
-  Different locations are known by different names
    -  Microglial in brain
    -  Kupffer in liver
    -  Histiocytes in connective tissue
    -  Osteoclasts in bone

Professional antigen presenting cells: constitutively activate MHC II

-  Macrophages
-  Dendritic cells
-  B lymphocytes

T-cell receptor: a surface bound receptor that is never secreted. Binds to processed antigens. CD3 sits adjacent to the alpha-beta complex.

Release of histamine and TNF is most associated with mast cells/basophils.

**Polymorphonuclear leukocytes** = {neutrophils, basophils, eosinophils}

3 fn of antibodies: ADCC, opsonisation, neutralisaton, (mast cell activation too?!)

cAMP is a second messenger that has the effect of activating protein A

Coupling of G-protein alpha subunit to an agonist occupied receptors initially leads to exchange of bound GDP with GTP

# GIFTs

## Arachidonic acid

```{r tol, echo=FALSE, out.width="500px"}
knitr::include_graphics("aa_pathway.png")
```

Phospholipids from cell membranes (when broken it is released into the cell triggering these pathways). Arachidonate metabolites $\in$ {Eicosanoids}. Glucocorticoids are steroids inhibiting COX-2 and includes cortisol, cortisone, prednisolone.

# CBL

## Drugs

### Acne

-  \textbf{Benzoyl peroxide} (topical) - applied topically is bactericidal against *P. acnes*.
-  \textbf{Doxycycline} (oral) $\in$ {Tetracyclines} which are broad spectrum antibiotics. Inhibits TLN by binding to bacterial 30S subunit.
-  \textbf{Roaccutane} (oral **isotretinoin** is also known as 13-cis-retinoic acid) $\in$ {Retinoid}. It is used as a treatment for severe acne however there are many side effects whilst it is a teratogen so must not get pregnant. Suppresses all the four aetiological factors responsible for acne

### Candidiasis

-  Fluconazole (oral) $\in$ {azole}. Azoles are antifungals that inhibit fungal cell membrane synthesis. Fluconazole inhibits lanosterol-C14-demethylase which inhibits lanosterol $\rightarrow$ ergosterol production. More expensive but only one oral dose is required.
-  Clotrimazole (topical) $\in$ {azole}. Inhibits ergosterol production as well. Is cheaper but it has to be administered for a week intra-vaginally.

### Transplant

-  Immunosuppresive therapy: Cyclosporin A, prednisolone, azathioprine $\in$ {immunsuppressive medication} given during RA, and kidney transplants. It gets incorporated into replicating DNA thus halting replication and it also inhibits purine synthesis. $\therefore$ it affects quickly proliferating cells such as B and T lymphocytes)
-  Ganciclovir is used to treat CMV. Inihibits viral DNA pol by being a competitive inhibitor of dGTP and is also a poor substrate for chain elongation.

### Vaccinations lul

## Pathogens

-  \textit{Aspergillus fumigatus}: infects immunocompromised hosts
-  *Staphyloccocus aureus*
-  *Candida albicans*

# Immunodeficiency diseases

**Cytomegaly virus (CMV)**

-  Part of herpes genus so virus remains latent and can reactivate if immune system is depressed
-  Common in AIDS patients, cancer patients, and organ transplant recipients
-  CMV has been found in most body fluids and thus can be spread in a variety of ways

**HIV**

-  Low CD4 T cell count, not undergone immunosuppressive therapy, not previously diagnosed with congenital immune system diseases, and has one or more AIDS defining conditions.
    -  CMV which disseminates. Usually just found in transplant patients.
    -  Tb is the leading cause of death in HIV(+) patients
    -  Pneumonia due to *Pneumocystis jiroveci* as this is a fungus and doesn't cause pneumonia in healthy individuals.
    -  Candidiasis which disseminates outside the usual locations of vagina and oral cavity.
    -  Cryptococcosis caused by yeast.
-  Saquinavir and ritonavir are antiretrovirals used to prevent and inhibit HIV. They are protease inhibitors.
-  Enfuvirtade inhibits viral cell penetration.

**Severe combined immunodeficiency (SCID)**

-  Diarrhoea, failure to thrive, and life-threatening infection

**Chronic granulomatous disease**

-  Phagocytes (neutrophils and macrophages) use a respiratory burst to destroy pathogens such as bacteria. This uses a ROS such as superoxide.
-  CGD - the phagocytes can't produce ROS meaning they can't kill bacteria. So patients have weakened immune system as their phagocytes are useless.
-  Leads to formation of granulomas in organs as they cannot phagocytose things.

**X-linked agammaglobulinaemia**

-  are more susceptible to pneumococcal infections

Selective IgA deficiency is the most common primary immune deficiency.

# Pathology

## Types of cells

-  Labile
-  Stable
-  Permanent: hypertrophy cannot hyperplasia

## Cellular changes

-  Atrophy: both $\Downarrow$Number and $\Downarrow$Size
-  Reversible: cellular swelling, fatty change, hyalinisation
-  Irreversible: rupture, mt damage, nuclear changes (e.g. karryohexis)

## Chronic inflamm

-  Tissue injury and repair co-exist
-  Mononuclear cells predominate: macro, lympho, plasma cells

## Granulomatous inflammation

-  \textit{Sometimes it happens that way, hyung.}
-  Things too big to be phagocytosed. E.g. Myobacterium tuberculosis, leprosy,...
-  T~H~1 detect antigen and secrete: IFN-$\gamma$ activating macrophages (epitheloid) leading to granuloma formation

## Wound healing

-  Regeneration: dead cells replaced by alive ones of same type (e.g. liver)
-  Repair: scar formation
-  Tissue repair: granulation tissue + wound healing
    -  Granulation tissue formation = {inflamm, clearance, ingrowth of granulation tissue}
    -  Wound healing = {first intention, second intention}
        1.  First intention: wounds with opposed edges (e.g. surgical incision). Not much fibrosis, not much scarring, scant inflamm, healing time is short. 
        2.  Second intention: fucked wound. Lots of fibrosis, scarring, extensive inflammation, takes a while.