### 2017-07-03 msk1.rmd
msk1

```
title: MSK I
author: Aaron Khoo
date: `r format(Sys.Date(%d %B %Y))`
```

# Readings

## Describing a fracture

1.  Describe the film: who what when where why
2.  What type of fracture? Complete = {transverse, oblique, spiral, comminuted}, Incomplete = {bowing, bucle, greenstick}, Salter-Harris
3.  Where is fracture = {diaphysis, metaphysis, epiphysis}
4.  Is it displaced? {angulation, translation, rotation}
5.  Something else happening too?

## Trauma

### Primary survey

-  Circulation (including haemorrhage control)
-  Airway Maintenance (cervical spine control)
-  Breathing and Ventilation
-  Circulation (reassess again - this is IMPORTANT)
-  Disability (check neurological status)
-  Exposure (undress patient completely - strip and flip, examining both sides of the patient) - cover as soon as practical to prevent hypothermia.

### Ix

-  FBC - Hb, Hct, Platelets
-  ABG with Lactate  - check for ventilatory failure and marker of shock (lactate)
-  Coag's - evidence of early trauma associated coagulopathy.
-  Blood alcohol
-  EUC
-  Group and Hold +/- X-Match as indicated.
-  ECG
-  CXR, PXR
-  eFAST (extended FAST ultrasound scan)
-  CT scans as indicated (generally in the obtunded major trauma patient a panscan is performed if the patient is haemodynamically stable)
-  Urinalysis