### 2017-07-04 msk1.rmd
Finished cbl

```
title: MSK I
author: Aaron Khoo
date: `r format(Sys.Date(%d %B %Y))`
output: pdf_document
```

# Readings

## Describing a fracture

1.  Describe the film: who what when where why
2.  What type of fracture? Complete = {transverse, oblique, spiral, comminuted}, Incomplete = {bowing, bucle, greenstick}, Salter-Harris
3.  Where is fracture = {diaphysis, metaphysis, epiphysis}
4.  Is it displaced? {angulation, translation, rotation}
5.  Something else happening too?

## Trauma

### Primary survey

-  Circulation (including haemorrhage control)
-  Airway Maintenance (cervical spine control)
-  Breathing and Ventilation
-  Circulation (reassess again - this is IMPORTANT; check for pulses, bleeding, cyanosis,...)
-  Disability (check neurological status)
-  Exposure (undress patient completely - strip and flip, examining both sides of the patient) - cover as soon as practical to prevent hypothermia.

### Ix

-  FBC - Hb, Hct, Platelets
-  ABG with Lactate  - check for ventilatory failure and marker of shock (lactate)
-  Coag's - evidence of early trauma associated coagulopathy.
-  Blood alcohol
-  EUC
-  Group and Hold +/- X-Match as indicated. (Checks blood type)
-  ECG
-  CXR, PXR
-  eFAST (extended FAST ultrasound scan)
-  CT scans as indicated (generally in the obtunded major trauma patient a panscan is performed if the patient is haemodynamically stable)
-  Urinalysis

## Principles of fracture Mx

### Initial

Maintain/improve:

-  Blood supply
-  Soft tissue health

Open fractures

-  Prevent infn: antibiotics immediately
-  Allow fracture to heal
-  Restore fn in the limb

### Non-operative therapy

-  Casting: contraindicated if = {no displacement, displacement exists but not relevant to fn outcome, reduction impossible, reduction cannot be maintained, fracture produced by traction forces}
-  Traction = {skin, skeletal}

### Surgical therapy

-  Kirschner wires
-  Plates = {buttress, compression, neutralisation, tension-band, bridge} and screws
-  External fixation: used with open fractures that have significant soft tissue disruption

# CBL Main case

## Trigger 1

58M 2 story fall, injuries right leg and arm, denies head or neck injury.

Immediate Mx: primary survey (see above).

Further info from pt:

-  HPC: What caused him to fall such as seizure or slip? How did he fall? Land on anything? Remember the whole thing?
-  Site, onset, character, radiation, associations, time course, exacerbating/relieving factors, severity (in-depth HPC later).
-  See if anything else is wrong. Distracting injury could obfuscate.
-  Who is next of kin?

After primary survey => secondary survey

-  Head to toe Ex
-  Look for fractures
-  Pulses, sensation, pain: neurovascularly intact
-  Auscultate for breath sounds, heart sounds
-  Palpate
-  FAST ultrasound: see if there is something bleeding
-  Log roll to turn them over and look for same stuff
    -  Push down c-spine to check if there is any pain

## Trigger 2

Pt tachycardic when he came in and hypotensive. Pulse gone down for a bit probably due to morphine getting rid of pain.

Imaging indicated: X-rays of right upper and lower limb. Can be difficult to assess for spinal injuries so it might be alright to image spine.

## Trigger 3

Fractures:

-  Complete spiral fracture to distal diaphysis R tibia
-  Comminuted fracture proximal diaphysis R fibula
-  Compound comminuted (through the skin) middle diaphysis right ulna
    -  Note: have to be methodical because if the ulna is fucked then the radius could be too

Fracture Mx:

-  Non-operative: closed reduction (pulling the bone back into alignment)
-  Operative: open reduction

Complications from open fracture:

-  Infn
-  Embolism

## Trigger 4

DDx post-op pain: 5W's of post-op fever = {wind [pneumonia], water [UTI], wound [infns especially after day 5], wonderdrugs [iatrogenic], walking [walking reduces DVT]}

-  Compartment syndrome
-  DVT
-  Infn
-  Nerve palsy

On Ex for new leg pain:

-  Inspection: signs on inflammation, cyanosis
-  Palpate: neurovascular, loss/decrease pulses, temperature, cap refill (comparing both sides)
-  Neuro assessment

## Trigger 5

Compartment system features:

-  Parasthesia
-  Tenderness
-  Increased pain from passive stretching (e.g. toes)