### 2017-07-10 msk2.rmd
added msk2

```
title: "MSK II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: pdf_document
```

# Bone

## Physiology

-  Matrix composed of 35% osteoid (mainly type I collagen) and 65% mineral component (mainly hydroxyapatite).
-  2 types of matrix = {woven, lamellar}
-  Cells = {osteoblasts, osteoclasts, osteocytes}
-  Development:
    -  Long bones from endochondral ossification
    -  Flat bones from intramembranous ossification
-  Remodelling:
    -  RANKL: stimulates RANK => increased osteoclast differentiation (stimulated by PTH)
    -  OPG: inhibits RANKL RANK signalling => decreased osteoclast differentiation

## Pathology

### Congenital

-  Dysostoses: defects in migration and condensation of mesenchyme
-  Dysplasia: global disorganisation of bone/cartilage
    -  Achondroplasia: most common AD disorder from g.o.f *FGFR3* mutation => retarded cartilage growth
    -  Thanatophoric dysplasia: lethal dwarfism also g.o.f *FGFR3*
    -  Osteogenesis imperfecta: mutation in *type I collagen* => dominant negative lof mutation => too little bone => fragility
    -  Osteopetrosis: reduced bone resorption resulting from impaired formation or fn of osteoclasts. AR mutations in *CA2* or *TCIRG1*

### Metabolic

-  Osteopenia
-  Osteoporosis: age related, reduced physical activity, genetic factors, calcium nutritional state, hormonal influences (e.g. estrogen deficiency).  
Clinical course can lead to lordosis/kyphoscoliosis/PE/pneumonia.
-  Rickets and osteomalacia: manifestations of vit D deficiency or abnormal metabolism

### Paget's

Paget disease is condition of increased but disordered and structurally unsound bone. Both genetic and environmental causes (usually *SQSTM1*). Hallmark is mosaic pattern of lamellar bone.

1.  Initial osteolytic stage
2.  Mixed osteoclastic/osteoblastic stage => predominance of osteoblastic activity
3.  Burned out quiescent osteosclerotic phase

### Fracture healing

-  Rupture of vessles results in haematoma surrounding area of bone injury
-  Clotted blood provides fibrin mesh, sealing off fracture site, and providing scaffold for inflamm cells/fibroblasts/angiogenesis.
-  Degradnulated platelets and inflamm cells release PDGF, TGF-$\beta$, FGF to activate osteoprogenitor cells.
-  End of first week: mass of uncalcified tissue called *soft callus* provides anchorage between ends of fractured bone.
-  After two weeks: soft callus => bony callus due to deposition of woven bone, hyaline/fibro-cartilage, and endochondral ossification.

### Osteomyelitis

Inflamm of bone and marrow, usually secondary to infn.

-  Pyogenic: usually bacteria, rarely fungi. Normally *Staph aureus*
    1.  Haematogenous spread
    2.  Extension from contiguous site
    3.  Direct implantation after compound fractures
-  Mycobacterial osteomyelitis

### Bone tumours

| Category          | Behavior  | Tumor Type                    | Common Locations                           | Age (yr) | Morphology                                                                                               |
|-------------------|-----------|-------------------------------|--------------------------------------------|----------|----------------------------------------------------------------------------------------------------------|
| Cartilage forming | Benign    | Osteochondroma                | Metaphysis of long bones                   | 10–30    | Bony excrescence with cartilage cap                                                                      |
| —                 | —         | Chondroma                     | Small bones of hands and feet              | 30–50    | Circumscribed hyaline cartilage nodule in medulla                                                        |
| —                 | Malignant | Chondrosarcoma (conventional) | Pelvis, shoulder                           | 40–60    | Extends from medulla through cortex into soft tissue, chondrocytes with increased cellularity and atypia |
| Bone forming      | Benign    | Osteoid osteoma               | Metaphysis of long bones                   | 10–20    | Cortical, interlacing microtrabeculae of woven bone                                                      |
| —                 | —         | Osteoblastoma                 | Vertebral column                           | 10–20    | Posterior elements of vertebra, histology similar to osteoid osteoma                                     |
| —                 | Malignant | Osteosarcoma                  | Metaphysis of distal femur, proximal tibia | 10–20    | Extends from medulla to lift periosteum, malignant cells producing woven bone                            |
| Unknown origin    | Benign    | Giant cell tumor              | Epiphysis of long bones                    | 20–40    | Destroys medulla and cortex, sheets of osteoclasts                                                       |
| —                 | —         | Aneurysmal bone cyst          | Proximal tibia, distal femur, vertebra     | 10–20    | Vertebral body, hemorrhagic spaces separated by cellular, fibrous septae                                 |
| —                 | Malignant | Ewing sarcoma                 | Diaphysis of long bones                    | 10–20    | Sheets of primitive small round cells                                                                    |

## Joints

-  Solid (nonsynovial: synarthroses): structural integrity
    -  Fibrous synarthroses: cranial structures and bonds between roots of teeth and jawbone
    -  Cartilaginous synarthroses (synchondroses): symphyses between sternum and ribs
-  Cavitated (synovial joints): have joint space for wide range of movement. Lined by type A synoviocytes (specialised macros with phago activity) and type B synoviocytes (sim to fibroblasts and synthesise hyaluronic acid). Hyaline cartilage serves as elastic shock absorber.

### Arthritis

-  Osteoarthritis: degeneration of articular cartilage leading to structural and fn failure of synovial joints.
-  Rheumatoid arthritis: chronic inflamm of autoimmune origin attacking the joints
-  Juvenile idiopathic arthritis: affects before age 16
-  Seronegative spondylarthropathies:
    -  Absence of rheumatoid factor
    -  Path changes in ligamentous attachments rather than synovium
    -  Involvement of sacroiliac joints $\pm$ others
    -  Association with HLA-B27
    -  Bony proliferation leading to ankylosis
-  Infectious arthritis
-  Suppurative arthritis: bacterial infns that cause acute suppurative usually enter joints from distant sites by haematogenous spread. Restricted ROM
-  Lyme disease: systemic infn by *Borella burgdorferi*
-  Crystal induced arthritis:
    -  Gout: urate crystal deposition in/around joints.
    -   Calcium Pyrophosphate Crystal Deposition Disease (Pseudogout)

### Tumours

Primary are rare.

-  Ganglion and Synovial Cysts: small (1–1.5 cm) cyst that is almost always located near a joint capsule or tendon sheath.
-  Tenosynovial Giant Cell Tumor

## Soft tissue tumours

Soft tissue refers to non-epithelial tissue excluding the skeleton, joints, central nervous system, hematopoietic and lymphoid tissues

### Tumours of adipose tissue

-  Lipoma: benign
-  Liposarcoma: malignant

### Fibrous tumours

-  Nodular fasciitis: self limiting fibroblastic and myofibroblastic proliferation occuring in upper extremities of young adults
-  Superificial fibromatosis: infiltrative proliferation causing local deformity but has innocuous clinical course
-  Deep fibromatosis: desmoid tumours. Large infiltrative masses that recur frequently but do not metastasise.

### Sk muscle tumours

-  Rhabdomyosarcoma: malignant mesenchymal tumour with sk muscle differentiation. Three main subtypes are recognized: alveolar [fusions of FOXO1 to PAX3/PAX7] (20%), embryonal (60%), and pleomorphic (20%).

### Smooth muscle tumour

-  Leiomyoma: benign most common in the uterus but can arise in any soft tissue site.
-  Leiomyosarcoma: They occur in adults and affect women more frequently than men.

### Tumours of uncertain origin

-  Synovial sarcoma
-  Undifferentiated pleomorphic sarcoma: ncludes malignant mesenchymal tumors with high-grade, pleomorphic cells that cannot be classified into another category by a combination of histomorphology, immunophenotype, ultrastructure, and genetics.