### 2017-07-12 histo_template.rmd
Added clean histo template but unusable in rmarkdown

---
title: "Histo of ____"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
header-includes:
  -  \usetheme{metropolis}
output:
  beamer_presentation:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("/home/khoo/Documents/uni_work/4_sem_1/7111/pracs/histo/week_")
# rename 's/test-this/REPLACESTRING/g' *
```

```{r image_loop2, results = 'asis'}
dickpics <- gtools::mixedsort(list.files("images/"))
for(i in 1:length(dickpics)){
  cat("### ", strsplit(dickpics[i], "\\.")[[1]][1], "\n", '\\includegraphics[width=\\linewidth,height=\\textheight,keepaspectratio]{{"', file.path("images", strsplit(dickpics[i], "\\.")[[1]][1]), '"}.PNG}', "\n", "\n", sep = "")
}
```

```{r image_loop2, results = 'asis'}
png_images <- gtools::mixedsort(list.files("images/"))  # Stores all the files in the dir images/
for(i in 1:length(png_images)){
    name <- strsplit(png_images[i], "\\.")[[1]][1]  # Splits up the string by "." and stores the first part (to get rid of .png)
    cat("\begin{frame}", "\n", 
  cat("### ", strsplit(dickpics[i], "\\.")[[1]][1], "\n", '\\includegraphics[width=\\linewidth,height=\\textheight,keepaspectratio]{{"', file.path("images", strsplit(dickpics[i], "\\.")[[1]][1]), '"}.PNG}', "\n", "\n", sep = "")
}
```