### 2017-07-12 latex_histo_template.tex

\begin{frame}

\begin{frame}
\frametitle{1 vincunt}
\includegraphics[width=\linewidth, height=\textheight, imagekeepaspectratio]{images/1 vincunt.png}
\end{frame}

\begin{frame}
\frametitle{2 full stops are aids}
\includegraphics[width=\linewidth, height=\textheight, imagekeepaspectratio]{images/2 full stops are aids.png}
\end{frame}

\begin{frame}
\frametitle{image}
\includegraphics[width=\linewidth, height=\textheight, imagekeepaspectratio]{images/image.png}
\end{frame}

\begin{frame}
\frametitle{rofllmao}
\includegraphics[width=\linewidth, height=\textheight, imagekeepaspectratio]{images/rofllmao.png}
\end{frame}

\end{frame}