### 2017-07-13 histo_template_latex_ver.rmd
added latex rmd flie that doesn't really work

---
title: "Histo of ____"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
header-includes:
  -  \usetheme{metropolis}
output:
  beamer_presentation:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("~/Documents/histo-test/")
# rename 's/test-this/REPLACESTRING/g' *
```

```{r image_loop, results = 'asis'}
image_name_ext <- gtools::mixedsort(list.files("images/"))  # Stores all the files in the dir images/ as a sorted character vector
image_name <- c()
for(i in 1:length(image_name_ext)){
    image_name[i] <- strsplit(image_name_ext[i], "\\.")[[1]][1]  # Splits up the string by "." and stores the first part (to get rid of file extension)

    # Outputting using LaTeX syntax to make new beamer slides, using double \\ 
    cat("\\begin{frame}", "\n", 
        "\\frametitle{", image_name[i], "}", "\n",
        "\\includegraphics[width=\\linewidth, height=\\textheight, imagekeepaspectratio]{",
        "images/", image_name_ext[i], "}", "\n",
        "\\end{frame}", "\n", "\n",
        sep = "")  # Make sure no unwanted spaces in between strings
}

# Obtaining second sub element of list sapply(strsplit(image_name_ext, "\\."), "[[", 1)
```