### 2017-07-13 histo_template.rmd
cleaned the histo_template

---
title: "Histo of ____"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
header-includes:
  -  \usetheme{metropolis}
output:
  beamer_presentation:
    keep_tex: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
setwd("~/Documents/histo-test/")
#setwd("/home/khoo/Documents/uni_work/4_sem_1/7111/pracs/histo/week_")
# rename 's/test-this/REPLACESTRING/g' *

### Requirements ###
# LaTeX engine with \usetheme{metropolis} (TeXlive-full?!)
# Directory called "images"
# ONLY image files in images that \includegraphics will allow for
# Pandoc/knitr (comes with RStudio by default)

### Notes ###
#    \includegraphics{{"../Current folder/1.This file"}.png}
# will not include the file name at the same time the quotations
# The extra curly brackets are to ensure we can choose it from a directory
# Below had to switch the ' and " around because LaTeX uses "
```

```{r image_loop2, results = 'asis'}
png_image_ext <- gtools::mixedsort(list.files("images/"))
png_image <- c()
image_ext <- c()
for(i in 1:length(png_image_ext)){
  png_image[i] <- strsplit(png_image_ext[i], "\\.")[[1]][1]  # removing the extension
  image_ext[i] <- strsplit(png_image_ext[i], "\\.")[[1]][2]  # keeping the extension

  cat("### ", png_image[i], "\n",  # name of the image
      "\\includegraphics[width=\\linewidth,height=\\textheight,keepaspectratio]{{",
      '"images/', png_image[i], '"}.', image_ext[i], "}", "\n", "\n", sep = "")
}
```