### 2017-07-16 msk3.rmd
added new msk week

---
title: "MSK III"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
---

# Readings

# Common paediatric hip diseases in primary care

## Developmental dysplasia of the hip

Most common in F firstborn infants. Inherited polygenic, acetabulum is shallow, and may contain pulvinar or fat.

Tests are Ortolani, Barlow, or Galeazzi signs.

## Slipped capital femoral epiphysis (SCFE)

Displacment of the capital femoral epiphysis from the rest of the femur through the cartilaginous growth plate known nas proximal femoral physis. Occurs  due to forces + acetabulum holding femoral head in place. This has sig proactive surgical methods to prevent severe hip arthritis. Risk factors = {obesity, male, african-american, pacific islander, endocrine disorders}

## Legg-Calve-Perthes disease

-  Synovitis (couple of weeks): reduced ROM
-  Osteonecrosis (6-12/12): contour of femoral head is maintained but part of it becomes ischaemic
-  Fragmentation phase (up to 3 years): pain, limping, reduced ROM
-  Revascularisation and regeneration (1-3 years): resorption of necrotic bone and dev of new bone. Permanent deformity can occur here.

# General approach to MSK

-  Inspection: SEADS = {swelling, erythema, atrophy, deformity, scars}
-  Palpation
-  ROM
-  Neurovascular status

## Treatment

-  Modification of activities
-  Ice
-  Compression
-  Elevation

# Main case

# Short case 1

# Short case 2