### 2017-07-18 msk3-gift.rmd
added gift

---
title: "Imaging in MSK"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
  word_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

Info from [this presentation](https://meds.queensu.ca/central/assets/modules/msk_imaging/attach/imaging_slides.pdf).

| Type | Shortened | Relative cost (2000 Canada) |
| -------- | -------- | -------- | -------- | -------- |
| Plain radiographs | X-ray | 1 (hand) |
| Computed tomography | CT | 8 (lumbar spine) |
| Magnetic resonance imaging | MRI | 18 (knee) |
| Ultrasound | USS | 2 (knee) | 
| Bone scan | | 7? |