### 2017-07-27 msk4.rmd
re-added

---
title: "MSK IV"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Readings

## Cancers

Described as a dull and aching pain in the location. Worse at night and sometimes increases with activity. Common symptoms also fever and night sweats. Many pt do not have any symptoms but notice a painless mass.

Primaries:

1.  Multiple myeloma: tumour of bone marrow.
2.  Osteosarcoma: most common in teenagers and children. Found femur, tibia, hip, shoulder.
3.  Ewings sarcoma: age between 5-20. Common locations are upper and lower leg, pelvis, upper arm, ribs. 
4.  Chondrosarcoma: 40-70 years of age. Found hip, pelvis, shoulder.

Common sites for secondaries: breast, lung, thyroid, renal, prostate

# Path POW

## Arthritis

If the PIP extended and DIP flexed: swan neck deformity

| OA | RA | Both |
| ------------ | ----------- | ----------- |
| Minimal joint inflammation | Morning joint stiffness lasts >1 hour | Affects small joints of the hands |
| Joint pain worsens after use | Presence of extra-articular manifestations  | Visible nodules on hands |
| | Abnormal levels of autoantibodies  | | 
| | Mostly symmetrical polyarthritis affecting young females | |

Radiological

-  Both: loss of jt space, malalignment of joints
-  OA: subchondral cysts and sclerosis, formation of bony outgrowths at the end of joints

# Main case

# Short case 1

## Trigger 1

Ulcer: local defect of tissue or organ caused by sloughing of inflammatory or necrotic material.

Pathogenesis of ulcer formation: possibly from kicking the bed, then repetitive trauma.

Factors preventing ulcer from healing (learn this):

-  Infn
-  Smoking
-  *Old* age
-  Heavy alcohol consumption
-  Medication: steroids, NSAIDs
-  Obesity
-  Repetitive trauma
-  Diabetic: due to decreased blood flow, increased risk for infn due to high sugar, neuropathy
-  Vascular disease (oxygen supply)
-  Poor hygiene + improper wound care
-  Malnutrition
-  Super rare genetic diseases

Wound present for long time:

-  Can affect gait/joints
-  Osteomyelitis
-  Scar
-  Could turn malignant (if wound not healing, always biopsy it)

# Short case 2