### 2017-07-27 msk4-nerve-quiz.rmd
quick save

---
title: "Nerves of upper limb"
author: "Aaron"
date: "27 July 2017"
output: beamer_presentation
header-includes: 
-  \usetheme{metropolis}
---

# Kahoot link

I don't know how exactly it works

https://play.kahoot.it/#/k/c06b4397-ac92-46c8-8d9a-1d74db17dd34

# What nerve supplies trapezius?

-  Spinal part of accessory nerve
-  Nerve to trapezius
-  Long thoracic nerve
-  Medial & lateral pectoral nerves

# What nerve supplies trapezius? (ans.)

-  \alert{Spinal part of accessory nerve}
-  Nerve to trapezius
-  Long thoracic nerve
-  Medial & lateral pectoral nerves

# A lesion to which of these nerves is most likely to cause "winged scapula"?

-  Dorsal scapula nerve
-  Nerve to subclavius
-  Long thoracic nerve
-  Medial and lateral pectoral nerves

# A lesion to which of these nerves is most likely to cause "winged scapula"? (ans.)

-  Dorsal scapula nerve (rhomboids, levator scapulae)
-  Nerve to subclavius (subclavius)
-  \alert{Long thoracic nerve (serratus anterior)}
-  Medial and lateral pectoral nerves (pec major + minor)

# Which nerve does NOT supply the rotator cuff muscles?

-  Suprascapular nerve
-  Subscapular nerve
-  Axillary nerve
-  Lower subscapular nerve

# Which nerve does NOT supply the rotator cuff muscles? (ans.)

-  Suprascapular nerve (supra/infraspinatus)
-  Subscapular nerve (subscapularis)
-  Axillary nerve (teres minor)
-  \alert{Lower subscapular nerve (teres major)}

# What muscle does the thoracodorsal nerve innervate?

-  Latissimus dorsi
-  Deltoid
-  Pectoralis major/minor
-  Trapezius

# What muscle does the thoracodorsal nerve innervate? (ans.)

-  \alert{Latissimus dorsi}
-  Deltoid
-  Pectoralis major/minor
-  Trapezius

#  Coracobrachialis, biceps brachii, and brachialis are innervated by which nerve?

-  Median nerve
-  Radial nerve
-  Nerve to the *brachi*
-  Musculocutaneous nerve

#  Coracobrachialis, biceps brachii, and brachialis are innervated by which nerve? (ans.)

-  Median nerve
-  Radial nerve
-  Nerve to the *brachii*
-  \alert{Musculocutaneous nerve}

#  Which nerve supplies the posterior group of arm muscles (triceps brachii, anconeus)?

-  Radial nerve
-  Median nerve
-  Nerve to the posterior arm muscles
-  Musculocutaneous nerve

#  Which nerve supplies the posterior group of arm muscles (triceps brachii, anconeus)? (ans.)

-  \alert{Radial nerve}
-  Median nerve
-  Nerve to the posterior arm muscles
-  Musculocutaneous nerve

#  In the superficial layer of anteromedial forearm muscles, what is NOT innervated by the median nerve? (bonus: what is it innervated by?)

-  Palmaris longus
-  Flexor carpi ulnaris
-  Pronator teres
-  Flexor carpi radialis

#  In the superficial layer of anteromedial forearm muscles, what is NOT innervated by the median nerve? (bonus: what is the other innervated by?) (ans.)

-  Palmaris longus
-  \alert{Flexor carpi ulnaris}
-  Pronator teres
-  Flexor carpi radialis

# What is innervated by the ulnar nerve? (fyi: anterior interosseous muscle that supplies a couple things here is a branch of the median nerve)

-  Flexor digitorum profundus medial part
-  Flexor digitorum profundus lateral part
-  Flexor digitorum superficialis
-  Flexor pollicis longus

# What is innervated by the ulnar nerve? (fyi: anterior interosseous muscle that supplies a couple things here is a branch of the median nerve) (ans.)

-  \alert{Flexor digitorum profundus medial part}
-  Flexor digitorum profundus lateral part
-  Flexor digitorum superficialis
-  Flexor pollicis longus

# What are *all* the muscles of the posterior forearm muscles innervated by (including their branches)?

-  Radial nerve (branches to deep branch of radial nerve then posterior interosseous nerve)
-  Median nerve (branches into anterior interosseous nerve)

# What are *all* the muscles of the posterior forearm muscles innervated by (including their branches)? (ans.)

-  \alert{Radial nerve (branches to deep branch of radial nerve then posterior interosseous nerve)}
-  Median nerve (branches into anterior interosseous nerve)

#  Brachial plexus is easy - Conrad 2017. Roots > Trunks > Divisions > Cords > Branches. Which of the 5 main nerve branches only comes from the lateral cord?

-  Axillary
-  Musculocutaneous
-  Radial
-  Ulnar
-  Median

#  Brachial plexus is easy - Conrad 2017. Roots > Trunks > Divisions > Cords > Branches. Which of the 5 main nerve branches only comes from the lateral cord? (ans.)

-  Axillary (posterior)
-  \alert{Musculocutaneous}
-  Radial (posterior)
-  Ulnar (medial)
-  Median (both medial and lateral cords)

#  Which of these intrinsic thumb muscles is innervated by the deep branch of the ulnar nerve?

-  Adductor pollicis
-  Abductor pollicis brevis (these and below all by recurrent branch of the median nerve)
-  Flexor pollicis brevis
-  Opponens pollicis

#  Which of these intrinsic thumb muscles is innervated by the deep branch of the ulnar nerve? (ans.)

-  \alert{Adductor pollicis}
-  Abductor pollicis brevis (these and below all by recurrent branch of the median nerve)
-  Flexor pollicis brevis
-  Opponens pollicis

#  Which nerve innervates the intrinsic little finger (not Peter Baelysh) muscles (abductor digiti minimi, flexor digiti minimi brevis, opponens digiti minimi)?

-  Recurrent branch of median nerve
-  Deep branch of median nerve
-  Deep branch of ulnar nerve
-  Recurrent branch of ulnar nerve

#  Which nerve innervates the intrinsic little finger (not Peter Baelysh) muscles (abductor digiti minimi, flexor digiti minimi brevis, opponens digiti minimi)? (ans.)

-  Recurrent branch of median nerve
-  Deep branch of median nerve
-  \alert{Deep branch of ulnar nerve}
-  Recurrent branch of ulnar nerve

#  Lumbricals and interossei generally supplied by?

-  Deep branch of ulnar nerve except lateral 2 lumbricals?
-  Deep branch of ulnar nerve except medial 2 lumbricals?

#  Lumbricals and interossei generally supplied by? (ans.)

-  alert{Deep branch of ulnar nerve except lateral 2 lumbricals?}
-  Deep branch of ulnar nerve except medial 2 lumbricals?

# Match the "hand" and nerve lesion together

-  Median nerve lesion: claw hand
-  Ulnar nerve lesion: claw hand
-  Diego Maradona: hand of god
-  Ulnar nerve lesion: hand of benediction

# Match the "hand" and nerve lesion together

-  Median nerve lesion: claw hand
-  alert{Ulnar nerve lesion: claw hand}
-  Diego Maradona: hand of god
-  Ulnar nerve lesion: hand of benediction