### 2017-08-03 neuro1.rmd
added sc

---
title: "Neuro I"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Readings

# Main case

## Trigger 1

Key observations:

-  Vitals
-  Oriented
-  Look at the wound: bleeding, discolouration, neurovascularly intact?
-  Any other injuries
-  Is he dangerous?

DRSABCDE: guy talking probably has patent airway and breathing. Circulation probably alright since his blood is trickling.

-  E = Exposure for secondary survey

MIST + AMPLE

M.  Mechanism of Injury
I.  Injuries Sustained
S.  Signs
T.  Treatment and Trends in the Vital Signs

A.  Allergies
M.  Medications
P.  Past Medical History
L.  Last Ate / LMP (in females)
E.  Events (This is usually the MIST)

Calming pt down:

-  "Why you so agitated?"
-  Just talk normally

Hx

-  HPC: how did you get injured?
-  Any numbness/tingling
-  What drugs are you on?
-  Which hand is dominant, when did you last eat?
-  Allergies

## Trigger 2

Info from physical

-  General inspection: mental state, conscious, oriented, body habitus, distress
-  Strength testing
-  Flexion/extension

## Trigger 3

Handover: include no signs of vascular compromise, mechanism of injury: fell through a window.

Numbness: damage to nerve which prevents sensation:

-  Compression
-  Laceration
-  Lesion
-  Demyelination
-  Ischaemia: pins and needles
-  Iatrogenic

Ix:

-  Xray: fracture/foreign body because worried about bleeding
-  FBC: check baseline haemoglobin in case he gets surgery

## Trigger 4

Delay before surgery: has to be nil-by-mouth, also theatre may be busy

Peripheral nerves can regenerate, CNS cannot. Process:

-  Needs to be realigned
-  Distal segment degrades, proximal grows through the sheath

## Trigger 5

Tinels sign: tap over nerve => if tingly/pins & needles => nerve irritation. Over time, expect the tingly feeling to go downwards more and more to show the nerve is healing.

Cock up splint to restore his finger flexion. Also so it doesn't atrophy.

# Short case 1

## Trigger 1

Anatomical site: bilateral is probably peripheral neuropathy.

Hx:

-  HPC: does it also occur in the hands? Also have to clarify if numb = tingling/weakness?

Ex: where exactly does it occur? Systemic or not?

DDx:

-  Diabetic neuropathy
-  Heavy metal toxicity
-  Alcohol
-  Nutritional disorders
-  GBS: often feel like they're walking on cotton

## Trigger 2

Other ANS involvement

-  Syncope/presyncope
-  Feeling hot/sweaty

Positive and negative symptoms:

-  Positive: might feel burning/tingling
-  Negative: loss of motor control, sensory, fine motor control

Pathology of peripheral neuropathies:

-  Axonal damage
-  Loss of myelination

Mononeuropathies:

-  Carpal tunnel syndrome
-  Tarsal tunnel
-  Bells palsy
-  Ulnar palsy: hitting funny bone

## Trigger 3

Ix: neuropathy screen (nerve conduction, biopsies, CSF (GBS)), bloods (FBC, CHEM20, antibodies (vasculitis)), fasting blood sugar

Mx for diabetes:

-  Diabetes: diet, metformin
-  Treatment of pain itself: antidepressents, anticompulsants (capsaicin, gabapentin)
-  Care of the foot:
-  Quit smoking

# Short case 2

## Trigger 1

Anatomical structures:

-  Carpal tunnel
-  Guillans tunnel
-  Unlikely cerebral hemispheres

## Trigger 2

DDx:

-  Carpal tunnel syndrome
-  Ganglion
-  Peripheral neuropathy
-  Injured nerve

## Trigger 3

Radiculopathy: other muscular and sensory deficits

Risk factors for carpal tunnel

-  Repeated wrist movements with force
-  Female: smaller carpal tunnel
-  40-60
-  Obesity

Diagnosis for carpal tunnel: hyperflexion test. Flex for 1 min and see if symptoms increase. Median nerve slow could be carpal tunnel syndrome.

## Trigger 4