### 2017-08-15 neuro3.rmd
Added neuro 3

---
title: "Module II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Readings

# Main case: online case with stroke

Types of speech deficits

-  Dysarthria: is a disorder of (motor) speech resulting from neurological injury of the motor component of the motor-speech
-  Dysphonia: hoarse voice
-  Dysphasia: disorder of language
    -  Expressive (Broca's): can't speak | damage in L frontal lobe reduced speech fluency with preserved comprehension. Fails to construct sentences well. Speech non-fluent.
    -  Reactive (Wernicke's): can't understand | L temporal-parietal damage leaves fluency of language but words are muddled. Cannot understand spoken or written words.

# Short case 1

# Short case 2