### 2017-08-24 neuro4.rmd
Added short cases, should try to start new sentence on new line from now on

---
title: "Neuro IV"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Readings

# Main case

## Trigger 1

Causes of irritability: high pitched cries

-  VINDICATE for the DDx

Parent Hx

-  Good Hx: associated Sx (vomiting, diarrhoea,...)
-  Outbreaks at the daycare
-  Any complications during birth (obstetric Hx)
-  Developmental Hx
-  How does Mad act usually?
-  How has her feeding changed
-  Nappy change rate
-  Rash?
-  Sleep habit change
-  Activity/interactivity level
-  Family Hx/sickness/who is living at home?

Assessment:

-  Paediatric assessment triangle
-  Primary survey (ABCDE)
-  Secondary survey
-  Ongoing assessment

## Trigger 2

Mx plan: address "what's wrong with her": you've brought her in early, but too early to figure out exact cause. Make sure to come back if it gets worse, come back even if not for vaccinations. Have to really figure out how the parents feel.

## Trigger 3

Physical signs

-  Poor cap refill in child

Meningitis: causes Hib, Neisseria, Strep

Under 3/12 & fever: send to ED

$\text{Septic screen} = \text{urine}+\text{bloods (FBC)}+\text{blood cultures} \pm \text{LP} $\pm$ \text{xray (if resp symptoms)}$

Meningitis:

|                   | Viral                      | Bacterial                                                            |   |   |
|-------------------|----------------------------|----------------------------------------------------------------------|---|---|
| Clinical symptoms | Headaches, increased ICP,… | Viral + confusion (much worse, progressesfaster as bacteria has pus) |   |   |
| Rate              | Way more common            | Less common                                                          |   |   |
| CSF               |                            |                                                                      |   |   |
|                   |                            |                                                                      |   |   |

Pathophysiology:

1.  Inflammation
2.  Infn => ventricles
3.  Necrosis, impaired CSF flow, CNS fn

Increases permeability of BBB

Strep pneumoniae is most common cause of MOPS = {meningitis, otitis media, pneumonia, sinusitis}

| Test                       | Bacterial                 | Viral                        |
|----------------------------|---------------------------|------------------------------|
| Opening pressure           | Elevated                  | Usually normal               |
| White blood cell count     | $\geq1000\text{ per }mm^{3}$| <100 per mm3            |  
| Cell differential          | Predominance of PMNs      | Predominance of lymphocytes |
| Protein                    | Mild to marked elevation  | Normal to elevated           |
| CSF-to-serum glucose ratio | Normal to marked decrease | Usually normal               |

## Trigger 4

Risk with LP: removal of CSF can suck brain through foramen magnum (coning)

# Short case 1

## Trigger 1

One eye: probably not occipital lobe as at optic chiasm, left optic tract gets info from both eyes (left visual field), sim for right.

Questions to ask:

-  Fam Hx
-  What was she doing at the time?
-  Any headaches when she does move her eye?
-  Associated symptoms: numbness, tingling, incontinence?
-  Nature of pain
-  Better or worse
-  Onset: there all day?
-  Allergies, drugs

## Trigger 2

Ex for pupilliary reflex

-  Direct
-  Consensual

## Trigger 3

Optic neuritis: inflammed myelin sheaths of optic nerve.
85% of cases permanent damage to optic nerve.
Tx: steroids for severe or chronic cases.

## Trigger 4

Causes of demyelinating disease: autoimmune, Vit B12,

Dx for MS: McDonald criteria: 2 diff lesions in 2 diff places and 2 diff times

# Short case 2

## Trigger 1

Peripheral nerves is site.

Breathless: up to 30% of GBS require ventilation.

## Trigger 2

Main cause is campylobacter causing GIT infns. 
Bacteria have LPS which are similar to myelin gangliocytes => creating molecular mimicry. 
Early lymphocytic infiltrates than macrophage mediated stripping of myelin. 
Segmental demyelination then infiltrates. 
Most repair normally.

| Mnemonic    | UMN                                 | LMN           |
|-------------|-------------------------------------|---------------|
| Strength    | $\Downarrow$                        | $\Downarrow$  |
| Tone        | $\Uparrow$                          | $\Downarrow$  |
| Others      | Superficial reflexes. Absent clonus | Fasiculations |
| Reflexes    | $\Uparrow$                          | $\Downarrow$  |
| Muscle mass | $\sim N$                            | $\Downarrow$  |
| Babinski    | $(+)$                               | $(-)$         |

Mx:

-  Plasmaphoresis or Ig but NOT both
-  20% mortality if ventilated
-  Long term complications DVT, paralysis, death, peripheral neuropathy
-  Recovery takes a *long* time: 6/12 to *years*