### 2017-09-05 neuro5.rmd
main case

---
title: "Neuro V"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Readings

# Main case

## Trigger 1

Symptoms due to old age: unilateral, stiff and weak, only 60 so probs not. Age related degeneration is typically bilateral, older,...

CNS

-  Motor power
-  Movement
-  Coordination: cerebellum

DDx:

-  Neuro: stroke, parkinsons, huntingtons, MS, GBS, MND (motor neuron disease), myasthenia gravis, cerebellar alcohol, vertigo, essential tremor, visual impairment that is not being picked up, dementia, peripheral neuropathy
-  Malig: primary or secondary tumours

Sensory for coordination:

-  Propioception
-  Visual
-  Vestibular apparatus

Hx taking:

-  Flesh out PC
    -  Onset: how did symptoms come in
    -  Fumbling at rest or moving
    -  Dizziness/falling over: vertigo, inner ear
    -  Visual fn: reading or not
    -  Headaches/faints: tumours
-  Anything else
-  PMHx
-  PSHx
-  FHx
-  Sexual Hx
-  Medications, allergies, adverse drugs
-  Social Hx: smoking, drinking, drugs, rest of it, how is it affecting life, dependents

## Trigger 2

Key symptoms

-  Slowing up: bradykinesia
-  Tremor: increases when trying to hurry or agitated
-  Mumbling
-  Episode of vertigo in the past: stemetil

Parkinsons: bradykesia, rigidity, tremor, postural instability

Basal ganglia: disinhibition and inhibition of differnet planned movements. Therefore has issues starting a movement.

| Extrapyramidal             | Corticospinal             | Cerebellar           |
|----------------------------|---------------------------|----------------------|
| Parkinsonism               | Increased tendon reflexes | Intention tremor     |
| Dystonia                   | Fasiculations             | Troubles with speech |
| Restlessness, urge to move | Increased tone            |                      |
| Gait: shuffling            |                           |                      |
| Lack of normal arm swing   |                           |                      |
| Walking heel to toe        |                           |                      |
| Cogwheel rigidity          |                           |                      |

## Trigger 3

Parkinsons: loss of 60-80 of dopaminergic neurons in pars compacta

Extrapyramidal disorders: anything disrupting the pathway. Lesions in basal ganglia, drugs

## Trigger 4

Side effects of levodopa:

-  Arrhythmias
-  Hypotension
-  Nausea
-  GI bleeding
-  Hallucinations
-  End of dose deterioration of fn
-  Oscillations
-  Dyskinesea

There are no disease modifying agents for parkinsons. These drugs only cure symptoms for a bit, disease will still progress.

## Trigger 5

On/off symptoms: try reduce dose but give more frequently so it still has effects without adverse effects

Natural Hx: loss of motor fn, dementia, death, autonomic effects

# Short case 1

# Short case 2