### 2017-09-05 neuro5.rmd
half added sc1

---
title: "Neuro V"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Readings

# Main case

## Trigger 1

Symptoms due to old age: unilateral, stiff and weak, only 60 so probs not. Age related degeneration is typically bilateral, older,...

CNS

-  Motor power
-  Movement
-  Coordination: cerebellum

DDx:

-  Neuro: stroke, parkinsons, huntingtons, MS, GBS, MND (motor neuron disease), myasthenia gravis, cerebellar alcohol, vertigo, essential tremor, visual impairment that is not being picked up, dementia, peripheral neuropathy
-  Malig: primary or secondary tumours

Sensory for coordination:

-  Propioception
-  Visual
-  Vestibular apparatus

Hx taking:

-  Flesh out PC
    -  Onset: how did symptoms come in
    -  Fumbling at rest or moving
    -  Dizziness/falling over: vertigo, inner ear
    -  Visual fn: reading or not
    -  Headaches/faints: tumours
-  Anything else
-  PMHx
-  PSHx
-  FHx
-  Sexual Hx
-  Medications, allergies, adverse drugs
-  Social Hx: smoking, drinking, drugs, rest of it, how is it affecting life, dependents

## Trigger 2

Key symptoms

-  Slowing up: bradykinesia
-  Tremor: increases when trying to hurry or agitated
-  Mumbling
-  Episode of vertigo in the past: stemetil

Parkinsons: bradykesia, rigidity, tremor, postural instability

Basal ganglia: disinhibition and inhibition of differnet planned movements. Therefore has issues starting a movement.

| Extrapyramidal             | Corticospinal             | Cerebellar           |
|----------------------------|---------------------------|----------------------|
| Parkinsonism               | Increased tendon reflexes | Intention tremor     |
| Dystonia                   | Fasiculations             | Troubles with speech |
| Restlessness, urge to move | Increased tone            |                      |
| Gait: shuffling            |                           |                      |
| Lack of normal arm swing   |                           |                      |
| Walking heel to toe        |                           |                      |
| Cogwheel rigidity          |                           |                      |

## Trigger 3

Parkinsons: loss of 60-80 of dopaminergic neurons in pars compacta

Extrapyramidal disorders: anything disrupting the pathway. Lesions in basal ganglia, drugs

## Trigger 4

Side effects of levodopa:

-  Arrhythmias
-  Hypotension
-  Nausea
-  GI bleeding
-  Hallucinations
-  End of dose deterioration of fn
-  Oscillations
-  Dyskinesea

There are no disease modifying agents for parkinsons. These drugs only cure symptoms for a bit, disease will still progress.

## Trigger 5

On/off symptoms: try reduce dose but give more frequently so it still has effects without adverse effects

Natural Hx: loss of motor fn, dementia, death, autonomic effects

# Short case 1

## Trigger 1

> 53M weakness and clumsiness of right hand 3/12. 6/12 began dropping things with R hand, gradual deterioration in handwriting, losing grip strength.

Anatomical sites (DDx for muscle weakness; unilateral/focal tends to be neurological))

-  Muscles:
    -  Primary myopathies: hereditoary/congenital (dystrophies), finalmm, metabolic
    -  Secondary myopathies: infn (trichanosis), conn tissue disease, thyroid/adrenal dysfn, paraneoplastic, drugs
    -  Trauma
-  NMJ: myasthenia gravis, lambert eaten, toxins such as snake venoms/ticks/pesticides
-  LMN: peripheral nerve entrapment, polyneuropathy, GBS
-  Spinal cord: radiculopathy from spinal stenosis, neoplasms, stuff damaging peripheral nerve roots, infn from CMV/lyme can damage anterior horn => LMN damage, MS
-  UMN: motor neuron diseases such as amyotrophic lateral sclerosis, 
-  Cerebral hemisphere: CVA (usually +sensory)
-  Systemic: Vit B12 deficiency, hypercalcaemia/hyperparathyroidism (+abd pain, excessive urination, renal stones)

Hx:

-  HPC
    -  Tell me more...
    -  What does weakness mean? Sometimes pt confuse weakness with fatigue, often part of systemic disorders, psych, MS, parkinsons. Pt might describe rigidity/incoordination as weakness, ask "what could you do that you no longer can" fishing for "lifting up milk",...
    -  Site
    -  Onset: insidious onset suggests degenerative or chronic (metabolic ,hereditary), intermittent (e.g. using it makes it worse suggests myasthenia gravis), particular positions making it bad then mechanical issue such as entrapment neuropathy, radiculopathy, localised vascular occlusion, particular times of day suggests ion channel disorder, random symptom timing with quick recovery suggests transient ischaemia of CNS
    -  Character, radiation
    -  Associated symptoms: 
        -  Atrophy/fasciculations suggest LMN
        -  Systemic symptoms (probs not unilateral arm) can be from endocrine (thyroid, diabetus)
        -  Neoplasias/infn
        -  *Sensory* such as parasthaesia, vision, vertigo suggests NS
        -  Change in character of voice, problems swallowing (pharyngeal muscles)
        -  Increased urinary urgency can suggest spinal damage (myelopathy)
    -  How is it affecting daily life, what does he think it is?
-  PMHx: can get clues for infn cause or neoplastic
-  SHx
-  FHx
-  Medications, allergies, ADR
-  Social Hx: drugs, alcohol, cigarettes

**finish up the DDx**

## Trigger 2

Cues:

-  Generally weak
-  No sensory or paraesthesia
-  Head, neck, or shoulder pain
-  No recent injuries
-  No pattern to weakness, not aware of stuff that makes it better or worse
-  No fatigueable element
-  No bladder or bowel disturbances, no double vision
-  No difficulties with swallowing or changes to voice/speech

Ex:

## References

-  [Good DDx for weakness](http://www.dartmouth.edu/~dons/part_2/chapter_12.html)
-  [AFP article: evaluation of pt with muscle weakness](http://www.aafp.org/afp/2005/0401/p1327.html)
-  [Medscape MND](http://emedicine.medscape.com/article/2111360-overview#a1)
-  [AFP article of MND](http://www.racgp.org.au/download/documents/AFP/2011/December/201112zoing.pdf)
-

# Short case 2