### 2017-09-11 neuro6.rmd
added week 10

---
title: "Neurology VI"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
urlcolor: blue
---

# Main case

# Short case 1

# Short case 2