### 2017-09-14 neuro6.rmd
sc

---
title: "Neurology VI"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Main case

## Trigger 1

DDx seizure/unconsciousness:

-  Vascular: CVA/TIA
-  Infn: encephalitis, meningitis, abscess
-  Trauma: head trauma
-  Neuro: MS
-  Medications/drugs/alcohol/sleep deprivation
-  Systemic: diabetes (hypo/hyper-glycaemia), hyperthyroidism, vit B12 deficiency, ion deficiency, young adult female = PREGNANT, dehydration (syncopal episode)

Initial assessment

-  DRSABC(DEFG = don't ever forget glucose)
-  MIST/AMPLE: mechanism of injury, injury sustained, vitals and signs/symptoms, treatments, / , allergies, medications, past illness, last meal, events related to injury

Seizures: inapproiate excitability/synchrony => increased bursts/decreases inhibition. Variety of genetic and structural defects that can contribute. Idiopathic epilepsy: bilateral generalised discharge.

Hx:

-  HPC: can she remember, onset: eating, how quickly LOC, drugs, previous episodes before, associated symptoms = {headaches, nausea, vomiting, aura}, history of head trauma, drugs/alcohol, [have to get collateral]
-  PMHx/SHx: genetic/congenital, Ca, mental disabilities, seizures before!, previous head trauma, neurosurgery
-  FHx
-  Repro Hx: sexually active
-  Allergies, medications, natural therapies + 
-  Social Hx: sleep, diet, if she lives alone, hobbies, drives for work,... % maybe don't need to find right *now* but do have to consider these

Seizures:

-  Focal: 
    -  Retained awareness
    -  Impaired awareness
-  General: tonic-clonic

Postictal state is the altered state of consciousness after an epileptic seizure

Ex: usually unremarkable for epilepsy, have to do if considering structural/infn

-  General inspection: checking for trauma
-  Neuro
-  CV

## Trigger 2

[Status epilepticus](https://lifeinthefastlane.com/ccc/status-epilepticus/):

-  Continuous seizure activity for 5 minutes or more without return of consciousness, or
-  Recurrent seizures (2 or more) without an intervening period of neurological recovery

Medications

-  Medazolam: benzodiazopine increases GABA so it is inhibitory
-  Phenytoin: anticompulsant that blocks V-gated Na channel

## Trigger 3

Hx:

-  [Febrile convulsions](http://www.rch.org.au/clinicalguide/guideline_index/Febrile_Convulsion/) are *common*
-  Focal seizures with impaired awareness leading to secondary generalised seizures: suggests structural abnormality
-  Absence seizures

Classification: still clasify with motor, non-motor, unclassified

-  Focal onset = {aware or impaired awareness}
-  Generalised = {tonic-clonic, absence, myoclonic (one muscle), clonic, tonic, atonic}
-  Unknown = {inadequate info, unusual presentation}

Further Hx:

-  Find the reason why she's had another seizure: triggered? forgot meds?
-  Fever, missing meals, pregnant,...

Ix:

-  EEG (+ sleep deprived to try provoke it)
-  MRI
-  Thyroid
-  Bloods: FBC

## Trigger 4

Temporal lobe: most common place for it to start. Temporal lobe epilepsy most common, mesial temporal sclerosis pathological finding. Fits with findings as she has focal => general stuff happening.

## Trigger 5

Medications: side effects generally drowsiness, learn which ones are for general vs focal epilepsy

-  Sodium valproate: increases production and inhibits breakdown of GABA | side effect = birth defects, tremor, thinning of hair
-  Carbamazepine: sodium channel blocker

Surgery: uncontrollable from medications, too severe. Can remove lobe or cut corpus callosum.

# Short case 1

## Trigger 1

DDx for 1st seizure:

-  Acute vascular event/trauma: haemorrhage
-  Infn: encephalitis > meningitis
-  Systemic: ionic imbalances, hypoglycaemia, hyperthyroidism
-  Drugs, medications
-  Tumour: glioma
-  Is she pregnant?
-  Epilepsy: might not have been apparent previously

## Trigger 2

Headache from space occupying lesion: worse on lying down (sleep/morning), coughing, sneezing. Can come from tumour, abscess,..

5 most common secondaries:

1.  Breast
2.  Lung
3.  Melanoma
4.  GIT/liver

Intrinsic: glioma,...

Extrinsic: meningioma, pituitary adenoma

Ix:

-  *Don't LP (yet): signs of increased ICP*
-  MRI
-  Bloods: blood glucose, FBCs (platelets, red cells), CHEM20 (ions, underlying kidney fn)
-  EEG 
-  PET later on to see if it was a secondary met

## Trigger 3

Mx:

1.  Surgical
2.  Radiotherapy
3.  Adjuvant chemotherapy: some passes BBB, some not

# Short case 2

## Trigger 1

Anatomical sites:

-  Incontinent of urine: could be generalised seizure
-  Right motor dysfn: left hemisphere

Hx:

-  HPC: ever had incontinence before, trauma, how was he right before, general health in the lead up, sitting or standing, 
    -  Systems neuro: changes in vision, headaches, dizziness/faint, nausea, memory problems, concentration, personality changes
-  PMHx/PSHx: had a brain surgery before, post CVA
-  Medications/allergies/adr/drugs/cigarettes/alcohol

## Trigger 2

1.  Can't maintain airway during a seizure: oxygen
2.  Stop seizure (status epilepticus): benzodiazepine
3.  IV access

## Trigger 3

Bloods:

-  FBC: RBCs, white cells
-  ESR/CRP: inflamm
-  CHEM20: underlying kidney/liver fn, specificaly Ca and Mg
-  Drug screen
-  ABG: due to resp symptoms

Ix:

-  Head CT
-  CXR
-  ECG

PDx: metastatic brain cancer

## Trigger 4

Further Ix:

-  Biopsy/chest CT to examine Ca better
-  PET

## Trigger 5

5 most common mets to brain:

-  Lung
-  Breast
-  Melanoma
-  Renal cell
-  Bowel