### 2017-09-18 git1.rmd
im so tired

---
title: "Module II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Main case

## Trigger 1

> 30s M distressed, bending over around abdomen, speaks no English

Anatomical systems causing abdo pain: (everything can)

-  CVS: ruptured AAA, MI
-  GIT: oesophagus, stomach, bile ducts, liver, large intestine, pancreas, spleen, small intestine, *peritoneum*
-  Neuro: radicular pain
-  MSK: rectus abd
-  Resp: 
-  Renal: nephrolithiasis, pyelo
-  Repro: testicular torsion
-  Endocrine:

DDx: (consider his age, is he YOPI?)

\includegraphics{images/git1-ddx-abdo.png}

GIT usually present with one of 4 signs:

1.  Abdo/chest pain
2.  Altered ingesting of food (from nausea, vomiting, dysphagia, odynophagia [painful swallowing], anorexia)
3.  ALtered bowel movements (diarrhoea or constipation)
4.  GI bleeding

Pathological processes: inflamm, neoplasm, ischaemia, obstruction, infn, haemorrhage, distension, chemical irritation (toxicity), movement, pressure

Visceral vs somatic pain:

-  Somatic = structures supplied by spinal nerves, includes the parietal pleura and peritoneum. Thus perforated duodenal ulcer and generalised peritonitis can cause diffuse abdo pain (T6-T12). 
-  Visceral: sympathetic component of the ANS (rami communicans => sympathetic chain => sphlancnic nerves [preganglionic efferent stimulatory signals; greater 59, lesser 1011, least 12, lumbar 12] => prevertebral vascular ganglia and plexues => postganglionic efferent fibres follow individual arteries to viscus

|                       | Somatic                                    | Visceral                                                       |
|-----------------------|--------------------------------------------|----------------------------------------------------------------|
| Origin                | Parietal peritoneum/adjacent musculature   | Viscera                                                        |
| Pain pathway          | Spinal nerves: dermatomes and myotomes     | Sympathetic NS                                                 |
| Perception            | Localised to region                        | Poorly localised or diffuse, often midline, distant  site      |
| Stimuli               | Inflamm, laceration, heat, chem irritation | Distension, ischaemia, inflamm                                 |
| Accompanying features | Local muscle spasm (guarding)              | Nausea, vomiting, tachycardia, $\Uparrow$BP, sweating, anxiety |

Location of abdo pain:

\includegraphics{images/git1-location-abdo.png}

Hx:

-  HPC: nature of the pain, how long it lasts, where is it located,...
-  GIT: $\Delta${Appetite, weight (malabsorption, malignancy, diabetes, thyroid, IBD)}, dysphagia/odynophagia, indigestion or heartburn, nausea or emesis, haematemesis (vomiting blood), abdo pain, jaundice, bowel habits, per rectal blood
-  PMHx, PSHx
-  Medications
-  Social history
-  FHx
-  Red flags in acute abdo pain: sudden onset pain, increasing severity, (pre-)syncope, emesis/haematemesis, abdo distension, pallour and diaphoresis, tachycardia or AF, hypotension, fever, rebound tenderness, decreased urine output, positive pregnancy test

## Trigger 2

> Afghan refugee, 3 years refugee camp with poor diet and unsanitary conditions, massive psychological and physical trauma, heavy smoker, does not drink alcohol. Abdo pain first occured 4/52 and recurring, always upper abdo, deep gnawing pain relieved by food, occasionally wakes at night, similar to something 3 years ago. Responsive to antacids, no Hx of diarrhoea or vomiting, black foul-smelling stool about 1/52, no Hx of NSAID usage.

Most common causes of recurrent epigastric pain: GORD and peptic ulcer disease (duodenal and gastric; can't separate based on Hx alone): gnawing and burning sensation, occuring after meals gastric (shortly after) and duodenal (2-3 hours afterwards)

Antacids: contain alkaline ions that neutralise the stomach acid

Normal: balance between gastric acid secretion and gastroduodenal defense (mucus and bicarbonate secretion, ion pumps to remove excess H+, tight junctions, cellular restitution, epithelial renewal). Patho: agressive factors such as NSAIDs or H pylori, alcohol, bile salts, acid, pepsin can alter this balance by allowing back diffusion of H+ => epithelial damage.

Gastric acid: important for digestion of proteins. From parietal cells which are regulated by

1.  Cephalic phase: sight, smell, taste, thoughts => medulla => vagal ouput => submucosal plexus => increases secretion from gastric glands and stomach motility
2.  Gastric phase: food reaching stomach => stretch and chemoreceptors => secretion of gastrin by G cells
3.  Intestinal phase: stimulation of intestinal receptors => secretion of duodenal hormones: *secretin* reducers gastric secretion, cholecystokinin inhibits gastric emptying

NSAIDs: Non-selective can damage the gastric and duodenal mucosal barrier and are associated with an increased risk of upper gastrointestinal ulceration, bleeding and perforation.

## Trigger 3

Ix:

-  Bloods: FBC, CHEM20, amylase, lipase
-  Endoscopy

Upper gastro intestinal endoscopy indicated: signs of bleeding, weight loss, vomiting, dysphagia, progression of symptoms, >55, FHx of gastrointestinal carcinoma. Helps rule out carcinoma, visualise ulcer, determine if bleeding

## Trigger 4

## Trigger 5

13C urea: high sensitivity and specificity, requires mass spec for Dx of *H. pylori*

# Short case 1

# Short case 2