### 2017-09-25 git2.rmd
readings lol

---
title: "Gastrointestinal II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Readings

[Explanation of diff git symptoms/signs](http://accessmedicine.mhmedical.com.ezproxy.library.uq.edu.au/content.aspx?bookid=1585&sectionid=96308009)

## Coeliacs

Coeliacs: gluten => gliadin (in the stomach; resists breakdown) => reaches the SI then binds to TFR (transferrin receptor; usually used to absorb iron, may be overexpressed) => transcytosed across enterocyte => lamina propia => gliadin is deamidated then phagocytosed by macrophages => macros are usually doing this for normal food anyway; abnormal reaction in coeliacs => Th2 (CD4+) => inflamm cytokines such as interferon-$\gamma$ and TNF-$\alpha$ => direct damage to epithelial cells in SI + killer CD8 t-cells stimulated => bad. 
Ab blood tests: anti-tTG (deamidates the gliadin) and anti-endomysial (EMAs). 
Symptoms = {children = [abd distension, failure to thrive, diarrhoea], adults = [chronic diarrhoea, bloating (distension),...]}. 
Tl;dr autoimmune.

## Diarrhoea

Diarrhoea: arises from four pathological processes

1.  Osmotic
2.  Secretory
3.  Altered motility
4.  Exudative/inflamm

Classified into chronic vs acute ($\leq2\text{week}$). 
Acute is usually infectious. 
Diarrhoea is bowel movements excessive in volume, frequency, or liquidity. 
It then interferes with nutrition. 
Causes can be infectiours or non-inf (i.e. drugs). Malabsorption: consistency reflects high nutrient and bacterial fermentation being bulky, pasty, or liquid, and tend to float in toilets (due to fats).

### Acute

Inflamm vs noninflamm

-  Noninflamm: watery, mild, self limited, Dx evaluation limited to pt with severe or persistent (>7 days) diarrhoea
-  Inflamm: blood or pus/fever, caused by invasive or toxin-producing bacterium, Dx requires routine stool bacterial cultures and testing as indicated for *C. difficile* toxin and ova and parasite

Mx:

\includegraphics{images/git2-acute-diarrhoea.png}

Rx:

-  Diet
-  Rehydration
-  Antidiarrhoeal agents: opioids, bismuth subsalicylate
-  Antibiotic therapy

# Main case

# Short case 1

# Short case 2