### 2017-09-26 git2.rmd
mc

---
title: "Gastrointestinal II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Readings

[Explanation of diff git symptoms/signs](http://accessmedicine.mhmedical.com.ezproxy.library.uq.edu.au/content.aspx?bookid=1585&sectionid=96308009)

## Coeliacs

Coeliacs: gluten => gliadin (in the stomach; resists breakdown) => reaches the SI then binds to TFR (transferrin receptor; usually used to absorb iron, may be overexpressed) => transcytosed across enterocyte => lamina propia => gliadin is deamidated then phagocytosed by macrophages => macros are usually doing this for normal food anyway; abnormal reaction in coeliacs => Th2 (CD4+) => inflamm cytokines such as interferon-$\gamma$ and TNF-$\alpha$ => direct damage to epithelial cells in SI + killer CD8 t-cells stimulated => bad. 
Ab blood tests: anti-tTG (deamidates the gliadin) and anti-endomysial (EMAs). 
Symptoms = {children = [abd distension, failure to thrive, diarrhoea], adults = [chronic diarrhoea, bloating (distension),...]}. 
Tl;dr autoimmune.

## Diarrhoea

Diarrhoea: arises from four pathological processes

1.  Osmotic
2.  Secretory
3.  Altered motility
4.  Exudative/inflamm

Classified into chronic vs acute ($\leq2\text{week}$). 
Acute is usually infectious. 
Diarrhoea is bowel movements excessive in volume, frequency, or liquidity. 
It then interferes with nutrition. 
Causes can be infectiours or non-inf (i.e. drugs). Malabsorption: consistency reflects high nutrient and bacterial fermentation being bulky, pasty, or liquid, and tend to float in toilets (due to fats).

### Acute

Inflamm vs noninflamm

-  Noninflamm: watery, mild, self limited, Dx evaluation limited to pt with severe or persistent (>7 days) diarrhoea
-  Inflamm: blood or pus/fever, caused by invasive or toxin-producing bacterium, Dx requires routine stool bacterial cultures and testing as indicated for *C. difficile* toxin and ova and parasite

Mx:

\includegraphics{images/git2-acute-diarrhoea.png}

Rx:

-  Diet
-  Rehydration
-  Antidiarrhoeal agents: opioids, bismuth subsalicylate
-  Antibiotic therapy

# Main case

## Trigger 1

DDx:

-  Motility: hyperthyroidism (hyper) or scleroderma (hypo)
-  Osmotic: improves with fasting
-  Exudative: damage to mucosa, e.g. shigella infn
-  Secretory: cholera

Most common cause in his age: infectious; chronic so thinking of some things like parasites, or immunocompromised

Fluids: 8-9L per day, only 1L is from intake whereas rest is from salivary glands,... 
SI = nutrients, LI = water

Hx:

-  HPC: onset, time course (how often, when, has it changed), consistency, how much he produces, how many wet nappies does he create, associated (systemicall unwell, any pain), was there a precipitating illness => antibiotics, who he comes into contact with (indigenous communities are a lot more communal)
-  PMHx: vaccinations

## Trigger 2

Poor diet: can by itself cause diarrhoea, e.g. full fat diet

## Trigger 3

Pale conjunctiva and nail beds: anaemia either from nutrition (Fe or Vit B12) or haemorrhage

Foul smelling, pale soft fluffy stool: malabsorptive. Smelly due to fermentation of the fats in it

Ix:

-  Bloods: FBC, ELFT, cultures (if he was febrile)
-  Stool sample

## Trigger 4

Cysts: not 100% specific or sensitive

Giardia:

-  Flagellated protozoa
-  Fecal-oral transmission
-  Proximal small bowel
-  Incredibly widespread: 70-80% of kids pass cysts but are asymptomatic

Anaemic: iron deficiency is microcytic, B12/folate are macrocytic |

-  Microcytic = {T = thalassaemia, I = iron deficiency, C = chronic disease, S = sideroblastic anaemia}
-  Macrocytic = {A = alcohol (cirrhosis), B = B12 deficiency, C = compensatory reticulocytosis (blood loss and hemolysis), D = drug/dysplasia, E = endocrine (hypothyroidism), F = folate deficiency/foetus (pregnancy)}

## Trigger 5

Metronidazole: inhibits nucleic acid synthesis so it affects both bacteria and protozoa

Long term consequences of parasitic infn:

-  Damage to gut mucosa
-  Can migrate to subarachnoid space => meningitis/encephalitis
-  Malnutrition
-  Failure to thrive

# Short case 1

# Short case 2