### 2017-09-28 git2.rmd
sc

---
title: "Gastrointestinal II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Readings

[Explanation of diff git symptoms/signs](http://accessmedicine.mhmedical.com.ezproxy.library.uq.edu.au/content.aspx?bookid=1585&sectionid=96308009)

## Coeliacs

Epidemiology: caucasians of european ancestry.

Coeliacs: gluten => gliadin (in the stomach; resists breakdown from gastric enzymes) => reaches the SI then binds to TFR (transferrin receptor; usually used to absorb iron, may be overexpressed) => transcytosed across enterocyte => lamina propia => gliadin is deamidated then phagocytosed by macrophages => macros are usually doing this for normal food anyway; abnormal reaction in coeliacs => Th2 (CD4+) => inflamm cytokines such as interferon-$\gamma$ and TNF-$\alpha$ => direct damage to epithelial cells in SI + killer CD8 t-cells stimulated => bad. 
Ab blood tests: anti-tTG (deamidates the gliadin) and anti-endomysial (EMAs). 
Symptoms = {children = [abd distension, failure to thrive, diarrhoea], adults = [chronic diarrhoea, bloating (distension),...]}. 
Tl;dr autoimmune.

## Diarrhoea

Diarrhoea: arises from four pathological processes

1.  Osmotic
2.  Secretory
3.  Altered motility
4.  Exudative/inflamm

Classified into chronic vs acute ($\leq2\text{week}$). 
Acute is usually infectious. 
Diarrhoea is bowel movements excessive in volume, frequency, or liquidity. 
It then interferes with nutrition. 
Causes can be infectiours or non-inf (i.e. drugs). Malabsorption: consistency reflects high nutrient and bacterial fermentation being bulky, pasty, or liquid, and tend to float in toilets (due to fats).

### Acute

Inflamm vs noninflamm

-  Noninflamm: watery, mild, self limited, Dx evaluation limited to pt with severe or persistent (>7 days) diarrhoea
-  Inflamm: blood or pus/fever, caused by invasive or toxin-producing bacterium, Dx requires routine stool bacterial cultures and testing as indicated for *C. difficile* toxin and ova and parasite

Mx:

\includegraphics{images/git2-acute-diarrhoea.png}

Rx:

-  Diet
-  Rehydration
-  Antidiarrhoeal agents: opioids, bismuth subsalicylate
-  Antibiotic therapy

# Main case

## Trigger 1

DDx:

-  Motility: hyperthyroidism (hyper) or scleroderma (hypo)
-  Osmotic: improves with fasting
-  Exudative: damage to mucosa, e.g. shigella infn
-  Secretory: cholera

Most common cause in his age: infectious; chronic so thinking of some things like parasites, or immunocompromised

Fluids: 8-9L per day, only 1L is from intake whereas rest is from salivary glands,... 
SI = nutrients, LI = water

Hx:

-  HPC: onset, time course (how often, when, has it changed), consistency, how much he produces, how many wet nappies does he create, associated (systemicall unwell, any pain), was there a precipitating illness => antibiotics, who he comes into contact with (indigenous communities are a lot more communal)
-  PMHx: vaccinations

## Trigger 2

Poor diet: can by itself cause diarrhoea, e.g. full fat diet

## Trigger 3

Pale conjunctiva and nail beds: anaemia either from nutrition (Fe or Vit B12) or haemorrhage

Foul smelling, pale soft fluffy stool: malabsorptive. Smelly due to fermentation of the fats in it

Ix:

-  Bloods: FBC, ELFT, cultures (if he was febrile)
-  Stool sample

## Trigger 4

Cysts: not 100% specific or sensitive

Giardia:

-  Flagellated protozoa
-  Fecal-oral transmission
-  Proximal small bowel
-  Incredibly widespread: 70-80% of kids pass cysts but are asymptomatic

Anaemic: iron deficiency is microcytic, B12/folate are macrocytic |

-  Microcytic = {T = thalassaemia, I = iron deficiency, C = chronic disease, S = sideroblastic anaemia}
-  Macrocytic = {A = alcohol (cirrhosis), B = B12 deficiency, C = compensatory reticulocytosis (blood loss and hemolysis), D = drug/dysplasia, E = endocrine (hypothyroidism), F = folate deficiency/foetus (pregnancy)}
-  RACH, BLOOD?

## Trigger 5

Metronidazole: inhibits nucleic acid synthesis so it affects both bacteria and protozoa

Long term consequences of parasitic infn:

-  Damage to gut mucosa
-  Can migrate to subarachnoid space => meningitis/encephalitis
-  Malnutrition
-  Failure to thrive

# Short case 1

## Trigger 1

DDx:

-  Repro/gynae: pregnancy, ovarian cysts/cancer, fibromatosis
-  GIT: hernia, malabsorption (lactose intolerance => osmotic diarrhoea), coeliacs, IBD, infective, peptic ulcer (sequelae include cancers)
-  Psych: depression, chronic stress
-  Systemic: anaemia (Fe deficiency can lead to diarrhoea)
-  Endocrine: hypothyroidism, diabetic ketoacidosis

Ex:

-  Inspection: scarring, swelling, asymmetry, skin changes, guarding, general discomfort, also inspect eyes for jaundice/anaemia
-  Palpation: masses (firm/fixed/flatulent), organs, belot (wiggle) the kideys, for pain
-  Auscultate
-  Percussion: if considering ascites
-  Rectal

## Trigger 2

Aphthous ulcers: benign mouth ulcers associated with systemic stuff like coeliacs

Ix:

-  Bloods: FBC (Hb = anaemia), thyroid fn test, HLA typing, B12, ELFT
-  Stool sample
-  Tissue typing: ?
-  Discuss possibility for endoscopy: skip lesion things like Crohns; wait for bloods  maybe

## Trigger 3

Ix: suggest of coeliacs disease (systemic autoimmune disease). IgA-tTG: deamidating gluten products

Further Ix:

-  Iron studies
-  Gastroscopy and biopsy
-  Diet

## Trigger 4

Coeliacs associated with FHx of thyroid issues and T1D, Trisomy-21, IBD, GORD

Complications: bloating, chronic diarrhoea, nutritional deficiency, increased risk of malignancy, higher motility

Mx:

-  Gluten free diet
-  Frequent monitoring of coeliacs and systemic nutrient levels

Gluten found in wheat, rye, HLA-DQ2 (95%) and HLA-DQ-8. Can present at any age and common (1%). Learn the histopath => GIT systems + diverse systemic symptoms

# Short case 2

## Trigger 1

DDx:

-  Hernias: read medscape article
-  Cancer

Hx:

-  Reducible/irreducible, strangulated/incarcerated (soft/hard)
-  Fluctuants: cyst can look like hernia
-  Worse

Lump:

-  Fixed is worse than mobile

## Trigger 2

Mx: surgery

Natural Hx: will only get worse

## Trigger 3

The hernia became strangulated => ischamia => bowel necrosis

Medical emergency

Vomiting:

-  Pain
-  Bowel obstruction