### 2017-10-03 git3.rmd
mc

---
title: "Gastrointestinal III"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Main case

## Trigger 1

DDx: upper vs lower, stuff higher up => melaena

-  Haemorrhoids: varicose vein of perianal region
-  Colitis: ischaemic, inflamm, infn
-  CRC
-  IBD: ulcerative colitis + crohn's disease (less common for blood)
-  Trauma
-  AVM
-  Diverticulitis
-  Fissure
-  Polyps/FAP
-  Arterioenteric fistula
-  Peptic ulcer

Hx:

-  HPC: onset, frequency, anything worse/better, change in diet, 
    -  Info about poo: offensive odour (fat/mucous), consistency (liquid/solid), watery, colour, volume
    -  How frequent is blood
-  Associated symptoms: pain, nausea/vomiting, systemic weight loss, fatigue, bleeding/bruising
-  PMHx/PSHx
-  FHx
-  Social Hx

## Trigger 2

Constipation: usually haemorrhoids or anal fissure especially in 27M. 27M shouldn't really be anaemic. Even in F shouldn't just say "menstruation"

Types of diarrhoea: (todo: whip up a table for these types)

-  Osmotic: retention of water in the bowel from accumulation of non-absorbable substances
-  Secretory: electrolyte imbalance => water diffuses into stool. Often from inflamm
-  Motility: changed motility from hyperthyroidism, IBS, partial gastrectomy
-  Inflammatory: colon is inflamed, typically causes blood diarrhoea, e.g. IBD
-  (Malabsorptive: stool will float due to fats)

This pt: exudative (blood and mucus) + inflammatory. Not malabsorptive as his stool

Mucous in GIT: secreted for protection. Pus (secreted from inflamed). IBS and IBD

Associated pain to rectal bleeding:

-  Rectal: fissures
-  Abdominal:
    -  Severe: ischaemic colitis
    -  Mild to moderate: enteritis

Loperamide: opiod receptor agonist that doesn't cross BBB, decreasing activity of myenteric plexus

## Trigger 3

Hospital admission: think holistically, this pt is young adult male who has a partner who can look after him, chance that he'll have to go to hospital at night => yes

## Trigger 4

Bloods: urea elevated due to dehydration; albumin elevated due to chronic bleed; Hb due to anaemia, but MCV normal so probs not iron deficiency(?); CRP+WCC due to inflamm

Sigmoidoscopy: uses shorter instrument which can be done a lot faster

Todo: Copy a table between UC and CD from Don's thing

Natural Hx IBD: extraintestinal symptoms such as eyes, => complications such as toxic megacolon,

Toxic megacolon: inflamm => dilation of bowel => increased permeability + perforation => $\Uparrow$ septic

Mx: principles start aggressive => less aggressive

-  Short term: steroids
-  Long term: immunosuppressants, resection for UC

## Trigger 5

IBD associated

-  Skin: psorosis
-  Jt: seroneg arthritis, ank spond
-  Liver: primary sclerosin cholangitis
-  Thromboses, metabolic bone diseases

Regular monitoring: if the disease is being managed, liver damage, females: pregnant and many drugs are teratogenic, common malignancy in immunosuppressed are skin, vaccines

# Short case 1

# Short case 2