### 2018-02-27 mfhd1-mc.rmd
added main case for week 4

---
title: mfhd1mc
output: ioslides_presentation
---

## Usual causes of abN by age

<a href="https://i.imgur.com/Tm5nXOb.png">
  <img src="https://imgur.com/Tm5nXObl.png" />
</a>

[source: uptodate](https://www-uptodate-com.ezproxy.library.uq.edu.au/contents/image?imageKey=OBGYN%2F61684&topicKey=OBGYN%2F3263&source=see_link)

## Menstrual cycle

<a href="https://i.imgur.com/d77w8GT.png">
  <img src="https://imgur.com/d77w8GTl.png" />
</a>

[source: osmosis](https://www.osmosis.org/learn/Menstrual_cycle)

## Menstrual cycle (hormones)

<a href="https://i.imgur.com/3NqfjYa.png">
  <img src="https://imgur.com/3NqfjYal.png" />
</a>

[source: wikipedia](https://upload.wikimedia.org/wikipedia/commons/f/f3/Figure_28_02_07.jpg)

## Vaginal (pelvic?) exam 4:38

<iframe width="756" height="426" src="https://www.youtube.com/embed/Z-O_JYtyQqE" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## Cervical screening

[go to table](http://www.cancerscreening.gov.au/internet/screening/publishing.nsf/Content/healthcare-providers)

## position of uterus (1)

<a href="https://i.imgur.com/7TUGzmM.png">
  <img src="https://imgur.com/7TUGzmMl.png" />
</a>

[dunno](http://y7177.com/sa/anato/uterus/anteversion,anteflexion2.jpg)

## position of uterus (2)

<a href="https://i.imgur.com/OTJVfb3.png">
  <img src="https://imgur.com/OTJVfb3l.png" />
</a>

*Moore's Clinical Anatomy*

## endometrial thickness

[normal](https://radiopaedia.org/articles/endometrial-thickness)

[abnormal+USS](https://radiopaedia.org/articles/abnormally-thickened-endometrium-differential-1)

## endometrial biopsy

<iframe width="560" height="315" src="https://www.youtube.com/embed/5ymMtSJ_qKI?start=51" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## the other one

<iframe width="560" height="315" src="https://www.youtube.com/embed/_Iv5IDmb5jQ?start=73" frameborder="0" allow="autoplay; encrypted-media" allowfullscreen></iframe>

## Histology

<a href="https://i.imgur.com/7wGCZKL.png">
  <img src="https://imgur.com/7wGCZKLl.png" />
</a>

[not the source but watch this video](https://www.youtube.com/watch?v=cfLBJs2kKBA)

## classification of endometrial hyperplasia

<a href="https://i.imgur.com/XxXhh9n.png">
  <img src="https://imgur.com/XxXhh9nl.png" />
</a>

[from medscape](https://www.medscape.com/viewarticle/507187)