### 2018-05-02 haem2.rmd
add haem2

---
title: "Haematology II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Main case

## Trigger 1

DDx:

-  Liver pathology can cause both thrombocytopaenia and loss of girth
-  Mets to bone marrow such as from lung
-  EBV
-  Lymphoma
-  Leukaemia
-  Multiple myeloma
-  Acquired haemophilia
-  Could be two separate pathologies

Ex:

-  Inspection: cachexia is hints of rapid weight loss (overall wt loss, loose skin, looks unhealthy)
-  Haem: lymph nodes, pallor, bony tenderness, scratch marks
-  Abdo: hepatosplenomegaly, jaundice/spider naevia/gynaecomastia/oedema for liver pathology
-  Resp: if worried about lung Ca

## Trigger 2

Top Dx

-  Lymphoma
-  Leukaemia
-  Metastatic Ca
-  Still consider HIV as generalised lymphadenopathy $\pm$ pain

Splenomegaly

-  Massive:
-  Large: malaria

## Trigger 3

Pancytopaenia

-  Lymphoma: pancytopaenia from bone marrow infiltration and crowding out so it cannot produce normal cells
-  Leukaemia: producing abnormal cells?
-  Aplastic anaemia: basically no more bone marrow left, cannot produce anything
-  Vitamin B12/folate deficiency
-  Medications

B symptoms: reserve for use when talking about lymphoma. Night sweats have to be drenching. If not quite B symptoms then it is called constitutional symptoms

## Trigger 5

-  Infiltration extramedullary haematopoeisis => hepatosplenomegaly => early satiety from pushing against it

# Short case 1

# Short case 2