### 2018-05-04 haem2.rmd
sc

---
title: "Haematology II"
author: "Aaron Khoo"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  pdf_document
documentclass: article
classoption: a4paper
fontsize: 11pt
urlcolor: blue
---

# Main case

## Trigger 1

DDx:

-  Liver pathology can cause both thrombocytopaenia and loss of girth
-  Mets to bone marrow such as from lung
-  EBV
-  Lymphoma
-  Leukaemia
-  Multiple myeloma
-  Acquired haemophilia
-  Could be two separate pathologies

Ex:

-  Inspection: cachexia is hints of rapid weight loss (overall wt loss, loose skin, looks unhealthy)
-  Haem: lymph nodes, pallor, bony tenderness, scratch marks
-  Abdo: hepatosplenomegaly, jaundice/spider naevia/gynaecomastia/oedema for liver pathology
-  Resp: if worried about lung Ca

## Trigger 2

Top Dx

-  Lymphoma
-  Leukaemia
-  Metastatic Ca
-  Still consider HIV as generalised lymphadenopathy $\pm$ pain

Splenomegaly

-  Massive:
-  Large: malaria

## Trigger 3

Pancytopaenia

-  Lymphoma: pancytopaenia from bone marrow infiltration and crowding out so it cannot produce normal cells
-  Leukaemia: producing abnormal cells?
-  Aplastic anaemia: basically no more bone marrow left, cannot produce anything
-  Vitamin B12/folate deficiency
-  Medications

B symptoms: reserve for use when talking about lymphoma. Night sweats have to be drenching. If not quite B symptoms then it is called constitutional symptoms

## Trigger 5

-  Infiltration extramedullary haematopoeisis => hepatosplenomegaly => early satiety from pushing against it

# Short case 1

Mech of fever: exogenous pyrogens => IL-1 and TNF => increased cyclooxygenase activity => $\Uparrow$ PG including PGE~2~ => reset temperature set point in hypothalamus

## Trigger 1

Fatigue in children + *infn*

-  Acute: *URTI*, sinus *infn*, hepatitis, chickenpox, measles, meningitis
-  More prolonged: *mononucleosis*, CMV, OSA
-  Chronic: tb, diabetes, *IBD*, CKD, thyroid disease, heart failure, poor nutrition, *cancer*, abscesses

Anaemia needs to be severe to cause persistent fatigue in children

## Trigger 2

"Growing pains": shouldn't affect the chest (sternum) but can affect children as young as 5

If generalised bone pain => really should consider leukaemia

## Trigger 3

Pale during fever because of vasodilation and less blood going to skin

Ix:

-  Good idea to do uric acid, LDH, Ca, Phos as a baseline in case tumour lysis syndrome

# Short case 2

## Trigger 1

Amoxiciliin + clavulanic acid: common for strep

DDx:

-  EBV
-  HIV
-  CMV
-  Acute hepatitis: abdo symptoms
-  Lymphoma: drenching night sweats
-  Withdrawal symptoms from alc/drugs can cause night sweats
-  Infective endocarditis

Histoplasmosis can come from bat bites

Pharyngitis is mostly viral

## Trigger 2

First line for tonsilitis is penicillin (amoxiclav)

Ix:

-  FBC
-  ELFT + creatinine = liver + kidney
-  CXR: tb in lungs
-  USS: spleen size, but it won't be diagnostic of prognostic so a bit useless..

## Trigger 3

Bloods:

-  Dehydration: osmolality high and urea high
-  ALT and AST: ALT = hepatocellular, AST = alcohol (sometimes)
-  Liver impairment