### 2018-05-10 haem3-sc1.rmd
sc1

---
title: Bad Formation (haem3sc1)
date: "10/05/2018"
author: Aaron Khoo
output:
  ioslides_presentation:
    incremental: true
    smaller: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Hx

Lorraine Cooper, is a new patient， 73F

<a href="https://i.imgur.com/QKrjYof.png">
  <img src="https://imgur.com/QKrjYofl.png" />
</a>

## Hx + Ex  (Trigger 1, 2)

You note on the chart that she is a 73 year old female with a local address. You greet Lorraine and ask her how you can help. She responds “To be perfectly honest with you doctor, I just want a second opinion. I have been **tired for some time now**, when I told my other doctor they did some tests which showed that my blood cells are no longer OK, especially my red cells– I think they call that anaemia. The doctor then said that my fatigue was due to my **anaemia** and this **should be investigated further with a number of tests**. Isn’t there something I could just take for this rather than more investigations? **I really don’t want any more poking and prodding**. You make a quick phone call to have Lorraine’s recent results sent to you and continue with your history.

You get some more background on Lorraine, and she tells you she is very well and in fact is **not on any medications**. Apart from some **osteoarthritis of her knees which she treats with hydrotherapy and physical exercise**, she has never had any serious illnesses. She has always been a **keen athlete** and attributes her good health to **regular physical activity**. She tells you she began **feeling fatigued over the course of the last year** and only brought it to her **usual doctor’s attention two weeks ago**. When she was told she was **anaemic**, Lorraine was partially surprised and didn’t realise anaemia could make her feel this tired. She describes **struggling to get out of bed in the morning and can feel a little stiff in her joints** but figured this was just aging. She has had to **give away bowls and tennis as she just doesn’t have the energy anymore**.

Her **examination is unremarkable**. Her **obs are stable** with a BP of 120/80mmHg and PR of 65. A haematological examination is unremarkable.

## Questions (Trigger 1)

-  **Given the story, what is the likely abnormality on Lorraine’s bloods?**
-  **What further aspects of history do you want to clarify?**
-  What is meant by fatigue? It can be used to describe
    -  Difficulty or inability to initiate activity (*subjective sense of weakness*)
    -  Reduced capacity to maintain activity (*easy fatigability*)
    -  Difficulty with concentration, memory, and emotional stability (*mental fatigue*)
-  Onset (ty uptodate)
    -  Acute fatigue: one month or less
    -  Subacute fatigue: between one and six months
    -  Chronic fatigue: over six months
        - Cause of chronic fatigue can be identified in approximately two-thirds of patients. Sometimes explained by overwork. Small number of cases are diagnosed with myalgic encephalomyelitis/chronic fatigue syndrome (ME/CFS) a disorder of unknown cause, but if the criteria is not met then it is known as idiopathic chronic fatigue. Both are diagnoses of exclusion.

## Fatigue

-  **When is fatigue pathological?**
-  Red flags: recent onset of fatigue in a previously well older patient, unintentional weight loss, abnormal bleeding, shortness of breath, unexplained lymphadenopathy, fever, recent onset or progression of cardio/gastro/neuro/rheum symptoms
-  **What are common causes for fatigue in this age group?**
-  **What are serious and not to be missed causes for fatigue in the elderly?**
-  The below is just general causes of fatigue (not spec to elderly)
-  *Probability diagnoses* include stress and anxiety, depression, viral/post-viral infection and sleep-related disorders
-  *Serious disorders* include malignancy, cardiac arrhythmias, cardiomyopathy, anaemia, bacterial infection, haemochromatosis, human immunodeficiency virus (HIV) infection and hepatitis C
-  *Common pitfalls* include coeliac disease, pregnancy, renal failure and metabolic disturbances. Medication-related side effects are another key consideration.

## Fatigue causes by system

-  **Cardiopulmonary**: congestive heart failure, chronic obstructive pulmonary disease, sleep apnoea.
-  **Endocrinologic/metabolic**: hypo/hyper-thyroidism, chronic renal disease, chronic hepatic disease, adrenal insufficiency, electrolyte abnormalities.
-  **Haematologic/neoplastic**: anaemia of diff causes, occult malignancy, myeloproliferative disorder, myelodysplastic syndromes, leukaemia/lymphoma, aplastic anaemia.
-  **Infectious**: Mononucleosis syndrome, viral hepatitis, HIV, subacute bacterial endocarditis, *tuberculosis*.
-  **Rheumatologic**: fibromyalgia, polymyalgia rheumatica, *systemic lupus erythematosus*, rheumatoid arthritis, Sjögren's syndrome, *sarcoidosis*.
-  **Psychological** (esp if prolonged fatigue): Depression, anxiety disorder, somatisation disorder.
-  **Neurological**: Multiple sclerosis.
-  **Medications**: Benzodiazepines, antidepressants, muscle relaxants, first-generation antihistamines, beta-blockers, opioids.
-  **Substance abuse**: EtOH, marijuana, opioids, cocaine/other stimulants.

## Bloods (Trigger 2)

<a href="https://i.imgur.com/WHAdzYZ.png">
  <img src="https://imgur.com/WHAdzYZl.png" />
</a>

## Simple blood cell lines

<a href="https://i.imgur.com/maYXUIT.png">
  <img src="https://imgur.com/maYXUITl.png" />
</a>

## Questions (Trigger 2)

-  Trigger 2 cont: You discuss with Lorraine that it would be worthwhile doing some further blood tests and considering referral to a haematologist. She is happy to oblige, anything to work out what's wrong.
-  **What pattern of abnormality is seen on the FBC?**
-  **What is the medical term used to describe this pattern of changes in her FBC and what are the causes?**
-  Pancytopenia (bicytopenia?): usually due to cancer, leukaemia, lymphoma, MM, TB, fungal, bone marrow aplasia (drug toxicities, nutrition, HIV, hepatitis, Parvovirus B19, autoimmune destruction)
-  **What other mechanisms can lead to a thrombocytopenia?**
-  *Decreased platelet production* (bone marrow failure/suppression, chronic EtOH, infn, MDS), *increased platelet consumption* (allo/auto-immune, DIC, sepsis, drug induced, infn, mechanical destruction by mech valves, HELLP, TTP/HUS), *sequestration/other* (chronic EtOH, hypersplenism, liver disease)
-  **What other possible causes of neutropenia are there?** [source](http://www.neutropenianet.org/wp-content/uploads/2013/07/publication-eval-treatment-of-scn.pdf)
-  Infn, drugs, autoimmune, nutritional, MDS, aplastic anaemia, cong bone marrow failure, neutropenia syndromes (GSD1b and other rare things)
-  **At what level of neutropenia does infection become a concern?**
-  0.2 - 0.5 increased risk of infn, < 0.2 risk of severe life threatening infn and opportunistic organisms

## Questions cont. (Trigger 2)

-  **What is the potential cause of high LDH level?**
-  Tissue damage: haemolytic anaemia, pernicious anaemia, infn, sepsis, AKI, acute liver disease, pacnreatitis, bone fractures, malignancy
-  **Is this anaemia proportionate to the degree of fatigue? How do the symptoms of anaemia evolve as the degree of anaemia increases?**
-  Age usually compounds fatigue, more anaemia = more fatigue, younger persons can tolerate more anaemia (no reputable source)
-  **What are your differential diagnoses?**
-  Pancytopenia: aplastic anaemia, mets to bone, MM, lymphoma, leukaemia, MDS, myelofibrosis, severe B12 folate deficiency, SLE, HIV/EBV, pernicious anaemia, medications, low dose arsenic poisoning, megaloblastic anaemia

## Further Ix (Trigger 3)

You repeat the blood tests which show a stable picture on the FBC, but specifically ask
for a reticulocyte count.

Additional results show that the reticulocyte count is 7 (10 - 100 x10^9^/L) indicating a
marrow cause for her anaemia.

Lorraine agrees to see a haematologist.

## Questions (Trigger 3)

-  **What is the significance of the abnormalities on the smear?**
-  **What are macrothrombocytes?**
-  Neutrophils: [normal](https://en.wikipedia.org/wiki/Neutrophil#/media/File:Neutrophils.jpg) vs [hypogranulated](https://en.wikipedia.org/wiki/Myelodysplastic_syndrome#/media/File:Hypogranular_neutrophil_with_a_pseudo-Pelger-Huet_nucleus_in_MDS.jpg)
-  [Oval macrocytes](https://i.imgur.com/IV5TEb7.png)
-  Macrothrombocytes: [smartsparrow](https://smartsparrow.global.ssl.fastly.net/repo/images/1d2066e379f598eeb9e9d4f6f101b90a.png)
-  **What are the causes of macrocytic anaemias?**
-  Macrocytic: increased reticulocytes (haemolytic anaemia, bone marrow recovery following chemo, recovery from bleed), megaloblastic (vit b12, folate, medications), multifactorial (EtOH, liver disease, HIV, MDS, aplastic anaemia, MM, preggers)
-  **How are the reticulocyte count and marrow involvement help in differentiating your hypotheses list?**
-  Aplastic anaemia, leukaemia, MDS, CHIP

## Trigger 4

Lorraine sees the haematologist who arranges for a bone marrow aspirate / trephine
which shows increased marrow cellularity with tri-lineage dysplasia. Iron stains show
normal iron stores but a small number of ring sideroblasts are present. The marrow has
been sent for cytogenetic testing.

The haematologist advises Lorraine that she has myelodysplastic syndrome and that
unfortunately this is a very serious incurable condition which will become progressively
worse. It is likely that the first intervention she will require will blood transfusions in the not too distant future.

## Questions (Trigger 4)

-  **What are the various forms of myelodysplasia?**
-  <a href="https://i.imgur.com/Nyrxp78.png">
  <img src="https://imgur.com/Nyrxp78l.png" />
</a>
-  [reading classification](https://academic.oup.com/view-large/35527240)
-  **What proportion of these are secondary to other causes? What are these?**
-  About 80% of people with MDS have primary MDS (no apparent risk factors). Secondary has poorer prognosis and risk factors are: previous chemo (alkylating agents/purine analogues), radiotherapy, tobacco smoking, benzene+derivatives, agricultural and industrial workers more likely
-  **At what point do blood transfusions become necessary? Are there complications associated with transfusion dependence?**
-  Chronic RBC transfusions are associated with chronic anaemia, leading to excess morbidity, and they cannot completely correct impaired QoL, may also be associated with iron overload => deleterious on organs (this is a bit disputed [MDS reading](https://academic.oup.com/annonc/article/25/suppl_3/iii57/1740877))
-  **At what level of platelets or neutrophils do complications usually arise?**

## Questions cont. (Trigger 4)

-  **What is the natural history of this disorder?**
-  Highly variable, survival ranges from a few weeks to several years. The median OS is 15-30 months and the risk of progression to AML is 25%-35% at 5 years. Bone marrow failure (infection and haemorrhage) is the leading cause of death, with more patients dying before overt AML has occurred
-  [Revised International Prognostic Scoring System (IPSS-R) for myelodysplastic syndromes](https://academic.oup.com/view-large/35527248) stratifies patients into five risk groups (v low, low, int, high, v high)
-  If she was at [high risk](https://oup.silverchair-cdn.com/oup/backfile/Content_public/Journal/annonc/25/suppl_3/10.1093_annonc_mdu180/3/mdu18002.png?Expires=1526047443&Signature=jd--FREEBGN0cNyVIGcYxEAWH01IU1W0N1Vx3Mur4v5rL~8BODfiZBoAlMgJAxu5cDYuOWgpT6ZeFkHU2jSy14oTgF7WDYZQPUpjyVUSSzDjtQXJO80-avbLy6YFDZZV2-D1aOusxmmaGXPXlgppRFqKdzPEHIKbXn60m4JJcwYxX-6Sph~B-3ArEekiCIIySUyD5hhRKR3LruN9FoLR9ir3s9WKb-6255hWMOzKaqXy1RrB-fMBB~dn750VLyn~~EuQQd31Ohu8UukEFNsGC5R3payOuEXIIAjr3btMhCw2yABTWmpS8zHqCLe9QAD3C4DmbGLl~kpQDCwUD436YA__&Key-Pair-Id=APKAIE5G5CRDK6RD3PGA)

## References

-  [RACGP fatigue](https://www.racgp.org.au/afp/2014/july/fatigue/)
-  [uptodate fatigue in adults](https://www-uptodate-com.ezproxy.library.uq.edu.au/contents/approach-to-the-adult-patient-with-fatigue)
-  [Thrombocytopenia aafp](https://www.aafp.org/afp/2012/0315/p612.html)
-  [Neutropenia by randos](http://www.neutropenianet.org/wp-content/uploads/2013/07/publication-eval-treatment-of-scn.pdf)
-  [Macrocytic anaemia uptodate](https://www-uptodate-com.ezproxy.library.uq.edu.au/contents/macrocytosis-macrocytic-anemia?topicRef=7133&source=see_link)
-  [MDS](https://academic.oup.com/annonc/article/25/suppl_3/iii57/1740877)
-  [Hematologylearning for some images](https://hematologylearning.weebly.com/white-blood-cell-inclusions-and-abnormalities.html)
-  [simpler MDS reading cancer.net](https://www.cancer.net/cancer-types/myelodysplastic-syndromes-mds/subtypes-and-classification)